### 2020-04-22 dev/fake_data.Rmd
First working version of fake data generation and data documentation

---
title: "Tests and development of fake data simulations"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
```

```{r}
fake_data <- fake_data_grein()
fake_data$breathing_data %>% ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile(width = 1, height = 0.5)
```

```{r}
fake_data$marker_data %>% filter(!(marker %in% c("hcq","az", "tocilizumab", "kaletra"))) %>% ggplot(aes(x = day, y = value, group = patient)) + geom_line() + facet_wrap(~marker, scales = "free")
```