### 2020-06-27 dev/hmm_grein.Rmd
Tweaks to m4

---
title: "Testing with Grein et al. data"
output: html_notebook
---

```{r setup}
library(rstan)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

devtools::load_all()

cache_dir <- here::here("local_temp_data")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}
```

## Load the data

```{r}

disease_levels <- c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")
breathing_levels <- c("Discharge", disease_levels, "Death")

data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  group_by(patient) %>%
  mutate(breathing_imputed = if_else(breathing == "ND", breathing[which.max(day[breathing %in% c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")])], breathing)) %>%
  ungroup() %>%
  mutate(
    breathing = factor(breathing, levels = breathing_levels, ordered = TRUE),
    breathing_imputed = factor(breathing, levels = breathing_levels, ordered = TRUE)
  )

series_data <- data_grein %>% transmute(.serie = as.integer(patient), .time = day + 1, .observed = breathing)

```

## Only increase/decrease

```{r}
m1_hidden_state_data <- tibble(id = factor(breathing_levels, levels = breathing_levels)) %>% mutate( corresponding_obs = id)

m1_observed_state_data <- tibble(id = factor(breathing_levels, levels = breathing_levels), is_noisy = FALSE)

recode_group <- function(rate_data, group_name) {
  is_column <- paste0("is_", group_name) 
  id_column <- paste0(group_name, "_id")

  rate_data %>% mutate(
    !!is_column := as.double(rate_group == group_name),
    !!id_column := factor(
      if_else(rate_group == group_name, as.integer(.rate_id) - min(as.integer(.rate_id[rate_group == group_name])), as.integer(0)), 
      levels = 0:(sum(rate_group == group_name) - 1), 
      ordered = TRUE)
  )
}

m1_rate_data <- rbind(
  tibble(.from = disease_levels, .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharge", rate_group = "improve_one"),
  tibble(.from = disease_levels[2 : length(disease_levels)], .to = disease_levels[1 : (length(disease_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = disease_levels[1 : (length(disease_levels) - 1)], .to = disease_levels[2 : length(disease_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(1:n()),
         .from = factor(.from, levels = breathing_levels),
         .to = factor(.to, levels = breathing_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one")

m1_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

m1_prior = brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id)
m1 <- brmshmmdata(m1_formula, series_data, 
                  m1_rate_data, m1_hidden_state_data, m1_initial_states, prior = m1_prior,
                  observed_state_data = m1_observed_state_data)

#make_stancode_hmm(m1)
#xx <- brms::make_standata(m1$formula, data = make_data_hmm(m1)$brmsdata)
m1_fit <- brmhmm(m1, cache_file = paste0(cache_dir,"/grein_m1.rds"))

summary(m1_fit$brmsfit)
```

```{r}
m1_epred_rect <- posterior_epred_rect(m1_fit)
m1_predicted_rect <- posterior_epred_to_predicted(m1_fit, m1_epred_rect)
m1_predicted <- posterior_rect_to_long(m1_fit, m1_predicted_rect)
```

```{r}
m1_predicted_df <- posterior_long_to_df(m1_fit$data, m1_predicted)

step_size <- 6
for(step in 1:ceiling(m1_fit$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m1_predicted_df, series_id, m1_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(m1_fit, m1_predicted_rect)
pp_check_transitions_direction(m1_fit, m1_predicted_rect)
```

```{r}
pp_check_transitions(m1_fit, m1_predicted_rect, states_from = 2:4)
pp_check_transitions(m1_fit, m1_predicted_rect, states_from = 5:7)

```

## Increase / decrease + noise

```{r}
m2_hidden_state_data <- rbind(
  tibble(base = disease_levels) %>% mutate(corresponding_obs = base) %>% 
    crossing(tibble(state_group = c("_improving","_worsening"))) %>% 
    mutate(id = factor(paste0(base, state_group))) %>%
    select(-base, -state_group),
  tibble(id = c("Death","Discharge")) %>% mutate(corresponding_obs = id)
)

m2_observed_state_data <- m1_observed_state_data

m2_rate_data <- rbind(
  tibble(.from = paste0(disease_levels, "_worsening"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA_improving", .to = "Discharge", rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels[2 : length(disease_levels)], "_improving"), 
             .to = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_improving"), 
             rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_worsening"), 
             .to = paste0(disease_levels[2 : length(disease_levels)], "_worsening"), 
             rate_group = "worsen_one"),
  tibble(.from = paste0(disease_levels,"_worsening"), 
             .to = paste0(disease_levels, "_improving"), 
             rate_group = "to_improving"),
  tibble(.from = paste0(disease_levels,"_improving"), 
             .to = paste0(disease_levels, "_worsening"), 
             rate_group = "to_worsening")
  ) %>%
  mutate(.rate_id = factor(1:n())) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  recode_group("to_improving") %>%
  recode_group("to_worsening")

m2_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")

m2_prior = m1_prior

m2_formula <- update.formula(m1_formula,
  ~ . + is_to_improving + is_to_worsening + (0 + is_to_improving | to_improving_id) + 
    (0 + is_to_worsening | to_worsening_id))

m2 <- brmshmmdata(m2_formula, series_data, 
                  m2_rate_data, m2_hidden_state_data, m2_initial_states, prior = m2_prior,
                  observed_state_data = m1_observed_state_data)

m2_fit <- brmhmm(m2, cache_file = paste0(cache_dir,"/grein_m2.rds"), control = list(adapt_delta = 0.9))

summary(m2_fit$brmsfit)
```

```{r}
m2_epred_rect <- posterior_epred_rect(m2_fit)
m2_predicted_rect <- posterior_epred_to_predicted(m2_fit, m2_epred_rect)
m2_predicted <- posterior_rect_to_long(m2_fit, m2_predicted_rect)
```

```{r}
m2_predicted_df <- posterior_long_to_df(m2_fit$data, m2_predicted)

step_size <- 6
for(step in 1:ceiling(m2_fit$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m2_predicted_df, series_id, m2_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(m2_fit, m2_predicted_rect)
pp_check_transitions_direction(m2_fit, m2_predicted_rect)
pp_check_transitions_direction(m2_fit, m2_predicted_rect, states_from = 2:7)
```

```{r}
pp_check_transitions(m2_fit, m2_predicted_rect, states_from = 2:4)
pp_check_transitions(m2_fit, m2_predicted_rect, states_from = 5:7)

```

## Improving/worsening + noise

```{r}
m3_hidden_state_data <- m1_hidden_state_data

m3_observed_state_data <- tibble(id =  breathing_levels, 
         is_noisy = c(FALSE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, FALSE),
         other_obs_1 = c(NA, disease_levels[2:6], disease_levels[1], NA), 
         other_obs_2 = c(NA, disease_levels[3:6], disease_levels[1:2], NA),
         other_obs_3 = c(NA, disease_levels[4:6], disease_levels[1:3], NA),
         other_obs_4 = c(NA, disease_levels[5:6], disease_levels[1:4], NA),
         other_obs_5 = c(NA, disease_levels[6], disease_levels[1:5], NA)

  ) %>% mutate(id = factor(id, levels = breathing_levels))

m3_rate_data <- m1_rate_data

m3_initial_states <- m1_initial_states

m3_prior = m1_prior

m3_formula <- m1_formula

m3 <- brmshmmdata(m3_formula, series_data, 
                  m3_rate_data, m3_hidden_state_data, m3_initial_states, prior = m3_prior,
                  observed_state_data = m3_observed_state_data)

m3_fit <- brmhmm(m3, cache_file = paste0(cache_dir,"/grein_m3.rds"), control = list(adapt_delta = 0.9))

summary(m3_fit)
summary(fit$brmsfit$fit, pars =  c("sensitivity", "other_observations_probs"), probs = c(0.025,0.975))$summary
```

```{r}
m3_epred_rect <- posterior_epred_rect(m3_fit)
m3_predicted_rect <- posterior_epred_to_predicted(m3_fit, m3_epred_rect)
m3_predicted <- posterior_rect_to_long(m3_fit, m3_predicted_rect)
```

```{r}
m3_predicted_df <- posterior_long_to_df(m3_fit$data, m3_predicted)

step_size <- 6
for(step in 1:ceiling(m3_fit$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m3_predicted_df, series_id, m3_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(m3_fit, m3_predicted_rect)
pp_check_transitions_direction(m3_fit, m3_predicted_rect)
pp_check_transitions_direction(m3_fit, m3_predicted_rect, states_from = 2:7)
```

```{r}
pp_check_transitions(m3_fit, m3_predicted_rect, states_from = 2:4)
pp_check_transitions(m3_fit, m3_predicted_rect, states_from = 5:7)

```

## Improving/worsening/stable

```{r}
m4_hidden_state_data <- rbind(
  tibble(base = disease_levels) %>% mutate(corresponding_obs = base) %>% 
    crossing(tibble(state_group = c("_improving", "_stable", "_worsening"))) %>% 
    mutate(id = factor(paste0(base, state_group))) %>%
    select(-base, -state_group),
  tibble(id = c("Death","Discharge")) %>% mutate(corresponding_obs = id)
)

m4_observed_state_data <- m1_observed_state_data

m4_rate_data <- rbind(
  tibble(.from = paste0(disease_levels, "_worsening"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA_improving", .to = "Discharge", rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels[2 : length(disease_levels)], "_improving"), 
             .to = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_improving"), 
             rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_worsening"), 
             .to = paste0(disease_levels[2 : length(disease_levels)], "_worsening"), 
             rate_group = "worsen_one"),
  tibble(.from = paste0(disease_levels,"_stable"), 
             .to = paste0(disease_levels, "_improving"), 
             rate_group = "to_improving"),
  tibble(.from = paste0(disease_levels,"_stable"), 
             .to = paste0(disease_levels, "_worsening"), 
             rate_group = "to_worsening"),
  tibble(.from = paste0(disease_levels,"_worsening"), 
             .to = paste0(disease_levels, "_stable"), 
             rate_group = "worsening_to_stable"),
  tibble(.from = paste0(disease_levels,"_improving"), 
             .to = paste0(disease_levels, "_stable"), 
             rate_group = "improving_to_stable"),
  tibble(.from = paste0(disease_levels,"_improving"), 
             .to = paste0(disease_levels, "_worsening"), 
             rate_group = "improving_to_worsening"),
  tibble(.from = paste0(disease_levels,"_worsening"), 
             .to = paste0(disease_levels, "_improving"), 
             rate_group = "worsening_to_improving")
  ) %>%
  mutate(.rate_id = factor(1:n())) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  recode_group("to_improving") %>%
  recode_group("to_worsening") %>%
  recode_group("improving_to_stable") %>%
  recode_group("worsening_to_stable") %>%
  recode_group("improving_to_worsening") %>%
  recode_group("worsening_to_improving")

m4_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")

m4_prior = m1_prior

m4_formula <- update.formula(m2_formula,
  ~ . + is_improving_to_stable + is_worsening_to_stable + is_improving_to_worsening + is_worsening_to_improving +
    (0 + is_improving_to_stable | improving_to_stable_id) + 
    (0 + is_worsening_to_stable | worsening_to_stable_id) +
    (0 + is_improving_to_worsening | worsening_to_improving_id) + 
    (0 + is_worsening_to_improving | improving_to_worsening_id) 
  )

m4 <- brmshmmdata(m4_formula, series_data, 
                  m4_rate_data, m4_hidden_state_data, m4_initial_states, prior = m4_prior,
                  observed_state_data = m1_observed_state_data)

m4_fit <- brmhmm(m4, cache_file = paste0(cache_dir,"/grein_m4.rds"), control = list(adapt_delta = 0.9), cores = 1)

summary(m4_fit$brmsfit)
```

```{r}
# m4_hidden_state_data <- rbind(
#   tibble(base = disease_levels) %>% mutate(corresponding_obs = base) %>% 
#     crossing(tibble(state_group = c("_improving", "_stable", "_worsening"))) %>% 
#     mutate(id = factor(paste0(base, state_group))) %>%
#     select(-base, -state_group),
#   tibble(id = c("Death","Discharge")) %>% mutate(corresponding_obs = id)
# )
# 
# m4_observed_state_data <- m1_observed_state_data
# 
# m4_rate_data <- rbind(
#   tibble(.from = paste0(disease_levels, "_worsening"), .to = "Death", rate_group = "death"),
#   tibble(.from = "AA_improving", .to = "Discharge", rate_group = "improve_one"),
#   tibble(.from = paste0(disease_levels[2 : length(disease_levels)], "_improving"), 
#              .to = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_improving"), 
#              rate_group = "improve_one"),
#   tibble(.from = paste0(disease_levels[1 : (length(disease_levels) - 1)],"_worsening"), 
#              .to = paste0(disease_levels[2 : length(disease_levels)], "_worsening"), 
#              rate_group = "worsen_one"),
#   tibble(.from = paste0(disease_levels,"_stable"), 
#              .to = paste0(disease_levels, "_improving"), 
#              rate_group = "to_improving"),
#   tibble(.from = paste0(disease_levels,"_stable"), 
#              .to = paste0(disease_levels, "_worsening"), 
#              rate_group = "to_worsening"),
#   tibble(.from = paste0(disease_levels,"_worsening"), 
#              .to = paste0(disease_levels, "_stable"), 
#              rate_group = "worsening_to_stable"),
#   tibble(.from = paste0(disease_levels,"_improving"), 
#              .to = paste0(disease_levels, "_stable"), 
#              rate_group = "improving_to_stable"),
#   tibble(.from = paste0(disease_levels,"_improving"), 
#              .to = paste0(disease_levels, "_worsening"), 
#              rate_group = "improving_to_worsening"),
#   tibble(.from = paste0(disease_levels,"_worsening"), 
#              .to = paste0(disease_levels, "_improving"), 
#              rate_group = "worsening_to_improving")
#   ) %>%
#   mutate(.rate_id = factor(1:n())) %>%
#   recode_group("death") %>%
#   recode_group("improve_one") %>%
#   recode_group("worsen_one") %>%
#   recode_group("to_improving") %>%
#   recode_group("to_worsening") %>%
#   recode_group("improving_to_stable") %>%
#   recode_group("worsening_to_stable")
#  
# 
# m4_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")
# 
# m4_prior = m1_prior
# 
# m4_formula <- update.formula(m2_formula,
#   ~ . + is_improving_to_stable + is_worsening_to_stable + (0 + is_improving_to_stable | improving_to_stable_id) + 
#     (0 + is_worsening_to_stable | worsening_to_stable_id))
#   
# m4 <- brmshmmdata(m4_formula, series_data, 
#                   m4_rate_data, m4_hidden_state_data, m4_initial_states, prior = m4_prior,
#                   observed_state_data = m1_observed_state_data)
# 
# m4_fit <- brmhmm(m4, cache_file = paste0(cache_dir,"/grein_m4.rds"), control = list(adapt_delta = 0.9))
# 
# summary(m4_fit$brmsfit)
```

```{r}
m4_epred_rect <- posterior_epred_rect(m4_fit)
m4_predicted_rect <- posterior_epred_to_predicted(m4_fit, m4_epred_rect)
m4_predicted <- posterior_rect_to_long(m4_fit, m4_predicted_rect)
```

```{r}
m4_predicted_df <- posterior_long_to_df(m4_fit$data, m4_predicted)

step_size <- 6
for(step in 1:ceiling(m4_fit$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m4_predicted_df, series_id, m4_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(m4_fit, m4_predicted_rect)
pp_check_transitions_direction(m4_fit, m4_predicted_rect)
pp_check_transitions_direction(m4_fit, m4_predicted_rect, states_from = 2:7)
```

```{r}
pp_check_transitions(m4_fit, m4_predicted_rect, states_from = 2:4)
pp_check_transitions(m4_fit, m4_predicted_rect, states_from = 5:7)

```