### 2020-07-09 dev/data_preprocessing.Rmd
Work on preprocessing

---
title: "Developing data load and processing"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
```

```{r}
raw_data_dir <- "H:/raw/"
if(!dir.exists(raw_data_dir)) {
  stop("Raw data dir not found")
}

raw_data_map <- readxl::read_excel(paste0(raw_data_dir,"mapping.xlsx"), sheet = "HospitalMapping") %>% filter(!is.na(File))
```

```{r}
file_relative <- raw_data_map$File[2]
hospital_id <- raw_data_map$HospitalID[2]

file <- paste0(raw_data_dir, file_relative)

patient_col_types <- c("text","numeric","text","date","date", #ID - Date of admission
                                   rep("text", 12), #Transferred from - Smoking
                                   rep("numeric", 5), # BMI - Albumin
                                   "text", "text", "date", "text", "text") #Discontinued medication - Transferred to

patient_data_raw <- readxl::read_excel(file, sheet = "Patients", range = "A2:AA100", na = "NA", col_types = patient_col_types)

logical_column <- function(x) {
  if(any( !is.na(x) && !(tolower(x) %in% c("yes","no"))) ) {
    stop("Invalid logical column contents")
  }
  tolower(x) == "yes"
}

outcome_labels <- c("Discharged","Hospitalized","Transferred","Death")

patient_data_typed <- patient_data_raw %>% 
  filter(!is.na(Age), !is.na(Sex)) %>%
  transmute(
    hospital_id = hospital_id, 
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = `Date of symptom onset`, 
    admission = `Date of admission`, 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = tolower(`Hypertension drugs`) != "no",
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = tolower(`Discontinued medication`) != "no",
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) == "no", lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) == "no", NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = `Date of last record`,
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels),
    transferred_to = `Transferred to`
  )
```