### 2020-08-17 dev/data_preprocessing.Rmd
Translation for preprocessing

---
title: "Developing data load and processing"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
```

```{r}
raw_data_dir <- "H:/raw/"
if(!dir.exists(raw_data_dir)) {
  stop("Raw data dir not found")
}

raw_data_map <- readxl::read_excel(paste0(raw_data_dir,"mapping.xlsx"), sheet = "HospitalMapping") %>% filter(!is.na(File))
```

Notes to resolve:
- Missing "Covid medication before hospitalization" in "Old" version.

```{r}
i <- 5
file_relative <- raw_data_map$File[i]
hospital_id <- raw_data_map$HospitalID[i]
lang <- raw_data_map$Lang[i]
file_version <- raw_data_map$Version[i]

file <- paste0(raw_data_dir, file_relative)

if(file_version == "From Sheet") {
  file_version <- readxl::read_excel(file, sheet = version_sheet(lang), range = "C8:C8", col_types = "text", col_names = "version")$version 
}

if(!grepl("^(Old|1\\.[2-4]\\.[0-9])$", file_version)) {
  print(file_version)
  stop("Invalid version")
}

if(file_version == "Old") {
  n_middle_text_columns <- 12
  last_column <- "Z"
} else {
  n_middle_text_columns <- 13
  last_column <- "AA"
}

patient_col_types <- c("text","numeric","text","date","date", #ID - Date of admission
                                   rep("text", n_middle_text_columns), #Transferred from - Smoking
                                   rep("numeric", 5), # BMI - Albumin
                                   "text", "text", "date", "text") #Discontinued medication - Outcome

patients_range <- paste0("A2:",last_column, "100")

patient_data_raw <- readxl::read_excel(file, sheet = patient_sheet(lang), range = patients_range, na = "NA", col_types = patient_col_types) %>% 
  translate_patient_columns(lang)

patient_data_raw <- patient_data_raw %>%
  mutate(`Outcome (Discharged/Hospitalized/Transferred/Death)` = translate_outcomes(`Outcome (Discharged/Hospitalized/Transferred/Death)`, lang))

if(file_version == "Old") {
  patient_data_raw <- patient_data_raw %>% mutate(`Covid Medications before admission` = NA_character_)
}
```

```{r}

patient_data_raw_usable <- patient_data_raw %>% filter(!is.na(Age), !is.na(Sex))

logical_column <- function(x) {
  if(any( !is.na(x) && !(tolower(x) %in% c(yes_variants, no_variants))) ) {
    stop("Invalid logical column contents")
  }
  tolower(x) %in% yes_variants
}

outcome_labels <- c("Discharged","Hospitalized","Transferred","Death")

# Special handling of badly filled outcomes
if(hospital_id == "QKuFp") {
  patient_data_raw_usable <- patient_data_raw_usable %>%
    mutate(`Outcome (Discharged/Hospitalized/Transferred/Death)` =
             if_else(grepl("^439[0-9]{2}$", `Outcome (Discharged/Hospitalized/Transferred/Death)`), 
                     "Hospitalized",
                     `Outcome (Discharged/Hospitalized/Transferred/Death)`))
}

patient_data_typed <- patient_data_raw_usable %>% 
  transmute(
    hospital_id = hospital_id, 
    patient_id = `Patient ID`,
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = lubridate::as_date(`Date of symptom onset`), 
    admission = lubridate::as_date(`Date of admission`), 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = !(tolower(`Hypertension drugs`) %in% no_variants),
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = !(tolower(`Discontinued medication`) %in% no_variants),
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) %in% no_variants, lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) %in% no_variants, NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = lubridate::as_date(`Date of last record`),
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels)
  )

invalid_outcome <- is.na(patient_data_raw_usable$`Outcome (Discharged/Hospitalized/Transferred/Death)`) != is.na(patient_data_typed$outcome)
if(any(invalid_outcome)) {
  print(patient_data_raw_usable %>% filter(invalid_outcome))
  stop("Invalid outcome")
}
```

```{r}
progression_data_raw <- readxl::read_excel(file, sheet = progression_sheet(lang), range = "A1:CU2539", na = c("NA",""), col_types = c(c("text","text", "skip", "text", "text"), rep("text", 94)), col_names = TRUE)

progression_data <- progression_data_raw %>% 
  translate_progression_columns(lang) %>%
  rename(patient_id = `Patient ID`, indicator = Indicator, first_day = First_Day, unit = Unit) %>%
  filter(!is.na(indicator)) %>%
  mutate(indicator = translate_markers(indicator, lang))

# Remove the "Don't edit examples" comment
progression_data$patient_id[2:9] <- NA_character_

# Spread Patient ID and Start date
last_patient_id <- progression_data$patient_id[1]
last_first_day <- progression_data$first_day[1]
for(i in 2:nrow(progression_data)) {
  if(!is.na(progression_data$patient_id[i])) {
    last_patient_id <- progression_data$patient_id[i]
    last_first_day <- progression_data$first_day[i]
  } else {
    progression_data$patient_id[i] <- last_patient_id
    progression_data$first_day[i] <- last_first_day
  }
}

progression_data <- progression_data %>% filter(patient_id %in% patient_data_typed$patient_id)

progression_data <- progression_data %>% 
  pivot_longer(starts_with("Day_"), names_to = "day", values_to = "value", names_prefix = "Day_") %>%
  mutate(day = as.integer(day))

progression_data <- progression_data %>% 
  filter(!is.na(indicator))

# Check dates
dates_data <- progression_data %>% 
  filter(indicator == "Date")
test_dates <- dates_data %>%
  mutate(value = as.integer(value), first_day = as.integer(first_day)) %>%
  filter(value != first_day + day - 1 )

if(nrow(test_dates) > 0) {
  stop("Dates check failed")
}

breathing_data <- progression_data %>% filter(indicator == "Breathing") %>%
  mutate(
    # Special cases
    value = case_when(value == "Oxygen/NIPPV" ~ "NIPPV",
                      TRUE ~ value),
    # Translate to factor
    breathing = factor(tolower(value), levels = tolower(c("AA","Oxygen", "NIPPV","MV","ECMO")),labels = c("AA","Oxygen", "NIPPV","MV","ECMO"), ordered = TRUE))

breathing_unrecognized <- breathing_data %>% filter(is.na(breathing) != is.na(value))
if(nrow(breathing_unrecognized) > 0) {
  print(unique(breathing_unrecognized$value))
  stop("Breathing NAs check failed")
}

breathing_data <- breathing_data %>% select(-value, -indicator, -unit)

correct_first_day_admission <- function(patient_data, progression_data) {
  nrow_before <- nrow(progression_data)
  progression_data <- progression_data %>% 
    inner_join(patient_data %>% select(patient_id, admission), by = "patient_id") %>%
    mutate(first_day = janitor::excel_numeric_to_date(as.numeric(first_day)),
           correction = as.integer(first_day - admission),
           day = day + correction - 1)

  if(nrow(progression_data) != nrow_before) {
    stop("Invalid join")
  }

  progression_data %>%
    select(-admission, -first_day, -correction)
}

breathing_data <- correct_first_day_admission(patient_data_typed, breathing_data) %>% filter(!is.na(breathing))

adverse_event_data <- progression_data %>% 
  filter(indicator == "Adverse events")

dummy_markers <- c("Drugs", "Other marker (optional)", "Other Covid drug", "Yet another drug")

marker_data <- progression_data %>% filter(!(indicator %in% c("Date", "Breathing", "Adverse events", dummy_markers))) %>% rename(marker = indicator)
marker_data <- correct_first_day_admission(patient_data_typed, marker_data)

# Correct Excel error
marker_data <- marker_data %>%
  mutate(marker = case_when(
    grepl("^Supp\\. O[0-9]*$", marker) ~ "Supp. O2",
    grepl("^SpO[0-9]*$", marker) ~ "SpO2",
    grepl("^IL-[0-9]*$", marker) ~ "IL-6",
    TRUE ~ marker))

if(any(is.na(marker_data$marker))) {
  stop("NA in markers")
}

marker_map <- c("PCR" = "pcr_value", 
                "Supp. O2" = "oxygen_flow",
                "SpO2" = "SpO2",
                "SpO2 native" = "SpO2_native",
                "Horowitz index" = "Horowitz_index",
                "PEEP" = "PEEP",
                "Ferritin" = "ferritin",
                "D-dimer" = "d_dimer",
                "CRP" = "CRP",
                "IL-6" = "IL_6",
                "PCT" = "procalcitonin",
                "Lymphocyte count" = "lymphocyte_count",
                "Leukocyte count" = "leukocyte_count",
                "Leukocytes" = "leukocyte_count",
                "Leukocyte" = "leukocyte_count",
                "Hydroxychloroquine" = "hcq",
                "Azithromycin" = "az",
                "Kaletra" = "kaletra",
                "Tocilizumab" = "tocilizumab",
                "Amoxicillin/Clavulanate" = "amoxiclav",
                "Piperacilin/tazobactam" = "piperacilin_tazobactam",
                "Clarithromycin" = "clarithromycin",
                "ofloxacin" = "ofloxacin",
                "ceftriaxone" = "ceftriaxone",
                "cefuroxim" = "cefuroxim",
                "cefotaxim" = "cefotaxim",
                "cefepim" = "cefepime",
                "cefepime" = "cefepime",
                "Vancomycin" = "vancomycin",
                "ampicillin/sulbactam" = "ampicillin_sulbactam",
                "zinkorot" = "zinc",
                "sulfamethoxazol" = "sulfamethoxazol",
                "trimethoprim" = "trimethoprim",
                "Meropenem" = "meropenem",
                "Fluconazol" = "fluconazol",
                "Covalescent plasma" = "convalescent_plasma",
                "Anidulafungin" = "anidulafungin",
                "Isoprinosine" = "isoprinosine"
                )
#Check NAs (TODO)
unrecgonized_markers <- marker_data %>% filter(!(tolower(marker) %in% tolower(names(marker_map))))

if(nrow(unrecgonized_markers) > 0) {
  print(unique(unrecgonized_markers$marker))
  print(unrecgonized_markers)
  stop("Unrecognized markers")
}

names(marker_map) <- tolower(names(marker_map))

marker_data <- marker_data %>% mutate(marker = marker_map[tolower(marker)])

# Special handling for PCR
pcr_data <- marker_data %>% filter(marker == "pcr_value")

pcr_values_positive <- c("pos","neg", "neg and pos", "susp.", "susp", "poz")
pcr_value_negative <- "neg"

wrong_pcr <- pcr_data %>% filter(!is.na(value),!(tolower(value) %in% pcr_values_positive), !tolower(value) %in% pcr_value_negative, !grepl("^[0-9]+$",value))
if(nrow(wrong_pcr) > 0) {
  print(wrong_pcr)
  stop("Wrong PCR")
}

pcr_data_long <- pcr_data %>% mutate(pcr_positive = as.numeric(tolower(value) %in% pcr_values_positive),
                    pcr_value = as.numeric(if_else(grepl("^[0-9]+$",value), value, NA_character_))) %>%
  select(-value, -marker) %>%
  pivot_longer(c("pcr_value","pcr_positive"), names_to = "marker", values_to = "value") %>%
  mutate(censored = "none")

# Handle NT for SpO2_native
marker_data <- marker_data %>% 
  mutate(value = if_else(marker == "SpO2_native" & tolower(value) == "nt", "< 92", value) )

# Some special cases
# TODO ask at hospital
if(hospital_id == "QKuFp") {
  marker_data <- marker_data %>% mutate(value = if_else(marker == "SpO2" & value == "NEG", NA_character_, value))
} else if(hospital_id == "YqNbe") {
  marker_data <- marker_data %>% mutate(value = if_else(marker %in% c("hcq","az") & value == "ano", ">100", value))
}

# Handle censoring on everything except PCR
marker_data <- marker_data %>% filter(marker != "pcr_value") %>%
  mutate(censored = case_when(startsWith(value, "<") ~ "right",
                              startsWith(value, ">") ~ "left",
                              TRUE ~ "none"),
         value_parsed = as.numeric(gsub("^[<>]? *", "", value)))

bad_parse_value <- marker_data %>% filter(is.na(value_parsed) & !is.na(value))
if(nrow(bad_parse_value) > 0) {
  print(bad_parse_value)
  stop("Bad parse value")
}

# Join with PCR data
marker_data <- marker_data %>%
  mutate(value = value_parsed) %>%
  select(-value_parsed) %>%
  rbind(pcr_data_long)

# Some special handling
if(hospital_id == "QKuFp") {
  marker_data <- marker_data %>% 
    mutate(unit = if_else(marker == "leukocyte_count" & "unit" == "unit", "10^9/l", unit))
}

```

```{r}
# TODO check units (and convert if necessary)
marker_data %>% group_by(marker)

```