### 2020-09-01 manuscript/descriptive.Rmd

---
title: "Descriptive stats and visuals"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
```

```{r}
nrow(data$patient_data)
data$breathing_data %>% group_by(patient_id) %>% summarise(n_days_data = n(), .groups = "drop") %>%
  ggplot(aes(x = n_days_data)) + geom_histogram(binwidth = 5)
```

```{r}
disease_levels <- c("Discharged", breathing_levels, "Death")
outcome_data <- data$patient_data %>% 
  filter(outcome %in% c("Discharged", "Death")) %>%
  transmute(patient_id = patient_id, 
            hospital_id = hospital_id,
            day = last_record,
            breathing = as.character(outcome))

data_for_plot <- data$breathing_data %>% 
  mutate(breathing = as.character(breathing)) %>%
  rbind(outcome_data) %>%
  mutate(breathing = factor(breathing, levels = disease_levels, ordered = TRUE))

data_for_plot %>% ggplot(aes(x = day, y = patient_id, fill = breathing)) + 
  geom_tile() + 
  scale_fill_viridis_d(direction = -1) +
  theme(axis.text.y = element_blank())
```
```{r}

```