### 2020-09-08 dev/raw_data_preprocessing.Rmd
First results on real data

---
title: "Developing data load and processing"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
```

TODO: code adverse events, code discontinued_medication_reason
TODO: check last day, add outcome to breathing data, possibly

```{r}
raw_data_dir <- "H:/raw/"
if(!dir.exists(raw_data_dir)) {
  stop("Raw data dir not found")
}

mapping_file <- paste0(raw_data_dir,"mapping.xlsx")

raw_data_map <- readxl::read_excel(mapping_file, sheet = "HospitalMapping") %>% filter(!is.na(File))
```

Notes to resolve:
- Missing "Covid medication before hospitalization" in "Old" version.

```{r}
hospital_data <- list()
for(i in 1:nrow(raw_data_map)) {
  file_relative <- raw_data_map$File[i]
  hospital_id <- raw_data_map$HospitalID[i]
  lang <- raw_data_map$Lang[i]
  file_version <- raw_data_map$Version[i]

  file <- paste0(raw_data_dir, file_relative)

  cat("Processing ", i, "-", hospital_id, "\n")

  hospital_data[[i]] <- read_raw_data(file, hospital_id, lang, file_version)
}

```

```{r}
complete_data <- join_hospital_data(hospital_data)
# TODO check:last day vs. breathing data
# MAx. day vs. read columns
differing_units <- complete_data$marker_data %>% group_by(marker) %>% filter(length(unique(unit)) > 1) %>% 
  group_by(marker, unit) %>% summarise(count = n(), .groups = "drop")

if(nrow(differing_units) > 0) {
  print(differing_units)
  print(complete_data$marker_data %>% group_by(marker) %>% filter(length(unique(unit)) > 1, !is.na(value)) %>% 
  ggplot(aes(x = value)) + geom_histogram() + facet_wrap(marker ~ unit, scales = "free")
  )

  stop("Differing units")
}

```

## Checking unit conversions

Check values for all markers with converted units (Note the 0 for Amoxiclav has been checked and is in the original data as well)

```{r}
for(conversion in unit_conversions) {
  for(marker_to_check in conversion$markers) {
    cat(" ==== ", marker_to_check, "=====\n")
    complete_data$marker_data %>% filter(marker == marker_to_check, !is.na(value)) %>% 
      pull(value) %>% unique() %>% print()
    print(
      complete_data$marker_data %>% filter(marker == marker_to_check, !is.na(value)) %>% 
        ggplot(aes(x = value)) + geom_histogram() + facet_wrap(marker ~ unit, scales = "free")
    )
  }
}
```
```{r}
complete_data$marker_data %>% filter(marker == "amoxiclav", value == 0)
```

```{r}
complete_data$marker_data %>% filter(marker == "CRP", unit == "mg/l") %>% pull(hospital_id) %>%   unique()

```

## Merging patients/wards

```{r}
patients_to_merge <- readxl::read_excel(mapping_file, sheet = "PatientsToMerge")
hospitals_to_merge <- readxl::read_excel(mapping_file, sheet = "HospitalsToMerge")
```

Check for candidates to merge

```{r}
for(i in 1:(length(hospital_data) - 1)) {
  for(j in (i + 1):length(hospital_data)) {
    #cat(i,",", j,"\n")

    candidate_overlap <-  check_patient_overlap(hospital_data[[i]], hospital_data[[j]], 
                                                patients_to_merge, hospitals_to_merge) %>% 
      select(hospital_id1, patient_id1, hospital_id2, patient_id2, 
             symptom_onset1, symptom_onset2, outcome1, outcome2, 
             BMI1, BMI2, last_record1, last_record2,
             transferred_from1, transferred_to2, transferred_from2, transferred_to1) 
    if(nrow(candidate_overlap) > 0) {
      # Workaround, issue https://github.com/rstudio/rstudio/issues/4439
      for(n in names(candidate_overlap)) {
        names(candidate_overlap[[n]]) <- NULL
      }
      print(candidate_overlap)
    }
  }
}
#  mutate(d = as.numeric(symptom_onset2 - symptom_onset1))
```

```{r}
complete_data <- merge_patients(complete_data, patients_to_merge)

complete_data$patient_data %>% filter(hospital_id == "Oriik", patient_id == 3)
```

```{r}
complete_data <- merge_hospitals(complete_data, hospitals_to_merge)

complete_data$patient_data %>% filter(hospital_id == "Oriik")
```

## Final processing steps

```{r}
no_data_pts <- 
  complete_data$breathing_data %>% group_by(hospital_id, patient_id) %>%
  summarise(count = n(), .groups = "drop") %>%
  filter(count < 1)

if(nrow(no_data_pts) > 0) {
  print(no_data_pts)
  stop("Some patients have no data")
}
```

```{r}
anonymized_data <- anonymize_for_analysis(complete_data)
```

```{r}
for(n in names(anonymized_data) ) {
  write_csv(anonymized_data[[n]], path = here::here("private_data", paste0(n,".csv")))
}
```