### 2020-09-16 manuscript/jm.Rmd

---
title: "Joint modelling"
output: html_notebook
---

```{r setup}
devtools::load_all()
library(tidyverse)
library(survival)
library(rstanarm)
library(tidybayes)

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- data$marker_data_wide

options(mc.cores = parallel::detectCores())

jm_hypothesis_res_list <- list()

cache_dir <- here::here("local_temp_data")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}

```

```{r}

data_surv <- wide %>%  
  filter(day > 0) %>%
  mutate(daym1 = day - 1,
    event_death = !is.na(breathing) & breathing == "Death",
    event_discharged = !is.na(breathing) & breathing == "Discharged",
  ) %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id"))  %>%
  arrange("patient_id", "day")

#TODO: impute the competing risk as censored long in the future
```

```{r}
data_d_dimer <- data_surv %>%
  group_by(patient_id) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
    select(event_death, event_discharged, 
         day, daym1, patient_id, hospital_id, 
         age, sex, took_hcq,
         d_dimer) %>%
  mutate(log_d_dimer = log(d_dimer + 1) - 4)

#data_d_dimer %>% filter(d_dimer == 0)
# data_d_dimer %>% group_by(sex, ever_hcq) %>% summarise(pts = length(unique(patient_id)), n_dimer = sum(!is.na(d_dimer)))
fit_d_dimer_age_sex <- stan_jm_with_cache(formulaLong = log_d_dimer ~ day + (1  | patient_id),
              dataLong = data_d_dimer %>% filter(!is.na(d_dimer)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex,
              dataEvent = data_d_dimer,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_d_dimer_age_sex.rds"))

fit_d_dimer_age_sex

#length(unique(data_d_dimer$patient_id))
# to_pred <- data_d_dimer %>% group_by(patient_id) %>% summarise(n_d_dimer = sum(!is.na(d_dimer))) %>%
#   filter(n_d_dimer > 1) %>% pull(patient_id)
#data_d_dimer %>% filter(patient_id == unique(data_d_dimer$patient_id)[31])

```

```{r}
jm_hypothesis_res_list[["d_dimer_age_sex"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$d_dimer_increases_death,
    fit_d_dimer_age_sex,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex")

jm_hypothesis_res_list[["d_dimer_age_sex"]]
```

```{r}
fit_d_dimer_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_d_dimer ~ day + (1  | patient_id),
              dataLong = data_d_dimer %>% filter(!is.na(d_dimer)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_d_dimer,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_d_dimer_age_sex_hcq.rds"))

fit_d_dimer_age_sex_hcq

#length(unique(data_d_dimer$patient_id))
# to_pred <- data_d_dimer %>% group_by(patient_id) %>% summarise(n_d_dimer = sum(!is.na(d_dimer))) %>%
#   filter(n_d_dimer > 1) %>% pull(patient_id)
#data_d_dimer %>% filter(patient_id == unique(data_d_dimer$patient_id)[31])

```

```{r}
jm_hypothesis_res_list[["d_dimer_age_sex_hcq"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$d_dimer_increases_death,
    fit_d_dimer_age_sex_hcq,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex, hcq")

jm_hypothesis_res_list[["d_dimer_age_sex_hcq"]]
```

```{r}
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[1:11])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[12:14])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[15:18])
plot(p1, plot_observed = TRUE)

p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[19:20])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[21:22])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[22:26])
plot(p1, plot_observed = TRUE)
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[27:29])
plot(p1, plot_observed = TRUE)                     
p1 <- posterior_traj(fit_d_dimer_age_sex_hcq, ids = unique(data_d_dimer$patient_id)[30:32])
plot(p1, plot_observed = TRUE)                     
```

```{r}
data_IL_6 <- data_surv %>%
  group_by(patient_id) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>% 
  select(event_death, event_discharged, 
         day, daym1, patient_id, hospital_id, 
         age, sex, took_hcq,
         IL_6) %>%
  mutate(log_IL_6 = log(IL_6))

fit_IL_6_age_sex <- stan_jm_with_cache(formulaLong = log_IL_6 ~ day + (1 + day | patient_id),
              dataLong = data_IL_6 %>% filter(!is.na(IL_6)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex,
              dataEvent = data_IL_6,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_IL_6_age_sex.rds"))

fit_IL_6_age_sex

```
```{r}
jm_hypothesis_res_list[["IL_6_age_sex"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$IL_6_increases_death,
    fit_IL_6_age_sex,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex")

jm_hypothesis_res_list[["IL_6_age_sex"]]

```

```{r}
fit_IL_6_age_sex_hcq <- stan_jm_with_cache(formulaLong = log_IL_6 ~ day + (1 + day | patient_id),
              dataLong = data_IL_6 %>% filter(!is.na(IL_6)),
              formulaEvent = Surv(daym1, day, event_death) ~ age + sex + took_hcq,
              dataEvent = data_IL_6,
              id_var = "patient_id",
              time_var = "day",
              cache_file = paste0(cache_dir, "/jm_IL_6_age_sex_hcq.rds"))

fit_IL_6_age_sex_hcq
```

```{r}
jm_hypothesis_res_list[["IL_6_age_sex_hcq"]] <-
  bayesian_hypothesis_res_from_jm(
    hypotheses$IL_6_increases_death,
    fit_IL_6_age_sex_hcq,
    coefficient_name = "Assoc|Long1|etavalue",
    adjusted  = "age, sex, hcq")

jm_hypothesis_res_list[["IL_6_age_sex_hcq"]]

```

```{r}
for(id in unique(data_IL_6$patient_id)) {
  p1 <- posterior_traj(fit_IL_6_age_sex_hcq, ids = id)
  print(plot(p1, plot_observed = TRUE)                     )
}

```

```{r}
jm_hypothesis_res_all <- do.call(rbind, jm_hypothesis_res_list)
write_csv(jm_hypothesis_res_all, path = here::here("manuscript", "jm_res.csv"))
```