### 2020-09-17 manuscript/competing_risks.Rmd
Workable draft

---
title: "Survival - frequentist"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
library(survival)

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- data$marker_data_wide

hypothesis_res_list <- list()

cache_dir <- here::here("local_temp_data")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}

```

# Competing risks death vs. discharged

```{r}
data_surv <- wide %>%  
  filter(day > 0) %>%
  mutate(daym1 = day - 1,
    status = fct_collapse(breathing, Hospitalized = breathing_levels) %>% factor(ordered = FALSE) %>% fct_relevel("Hospitalized", "Discharged", "Death") %>% fct_explicit_na("Hospitalized")
  ) %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id"))

f_competing_hcq <- Surv(daym1, day,status) ~ age + sex + took_hcq

#survcheck(f_competing_hcq, data = data_surv, id = patient_id)
fit_competing_hcq <- coxph(f_competing_hcq, data = data_surv, id = patient_id)
fit_competing_hcq

hypothesis_res_list[["competing_hcq"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq,
                                        coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex"),
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq,
                                        coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex")
  )

hypothesis_res_list[["competing_hcq"]]
```

```{r}
f_competing_hcq_az <- update.formula(f_competing_hcq, . ~ . + took_az)
fit_competing_hcq_az <- coxph(f_competing_hcq_az, data = data_surv, id = patient_id)

fit_competing_hcq_az

hypothesis_res_list[["competing_hcq_az"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az,
                                        coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, az"),
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az,
                                        coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, az"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_reduces_death, fit_competing_hcq_az,
                                        coefficient_name = "took_azTRUE", transition = "1:3", adjusted = "age, sex, hcq"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_increases_discharged, fit_competing_hcq_az,
                                        coefficient_name = "took_azTRUE", transition = "1:2", adjusted = "age, sex, hcq")
  )
hypothesis_res_list[["competing_hcq_az"]]
```

```{r}
f_competing_hcq_az_admitted <- update.formula(f_competing_hcq_az, . ~ . + admitted_for_covid)
fit_competing_hcq_az_admitted <- coxph(f_competing_hcq_az_admitted, data = data_surv, id = patient_id)

fit_competing_hcq_az_admitted

hypothesis_res_list[["competing_hcq_az_admitted"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az_admitted,
                                        coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, admitted, az"),
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az_admitted,
                                        coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, admitted, az"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_reduces_death, fit_competing_hcq_az_admitted,
                                        coefficient_name = "took_azTRUE", transition = "1:3", adjusted = "age, sex, admitted, hcq"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_increases_discharged, fit_competing_hcq_az_admitted,
                                        coefficient_name = "took_azTRUE", transition = "1:2", adjusted = "age, sex, admitted, hcq")
  )
hypothesis_res_list[["competing_hcq_az_admitted"]]
```

```{r}
f_competing_hcq_az_hospital <- update.formula(f_competing_hcq_az, . ~ . + hospital_id)
fit_competing_hcq_az_hospital <- coxph(f_competing_hcq_az_hospital, data = data_surv, id = patient_id)
fit_competing_hcq_az_hospital

hypothesis_res_list[["competing_hcq_az_hospital"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az_hospital,
                                        coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, az, hospital"),
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az_hospital,
                                        coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, az, hospital"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_reduces_death, fit_competing_hcq_az_hospital,
                                        coefficient_name = "took_azTRUE", transition = "1:3", adjusted = "age, sex, hcq, hospital"),
  frequentist_hypothesis_res_from_coxph(hypotheses$az_increases_discharged, fit_competing_hcq_az_hospital,
                                        coefficient_name = "took_azTRUE", transition = "1:2", adjusted = "age, sex, hcq, hospital")
  )
hypothesis_res_list[["competing_hcq_az_hospital"]]

```
```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = here::here("manuscript", "competing_risks_res.csv"))
```

```{r}
f_competing_hcq_az_comorbidities_sum <- update.formula(f_competing_hcq_az, . ~ . + comorbidities_sum)
fit_competing_hcq_az_comorbidities_sum <- coxph(f_competing_hcq_az_comorbidities_sum, data = data_surv, id = patient_id)
fit_competing_hcq_az_comorbidities_sum

hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_reduces_death, fit_competing_hcq_az_comorbidities_sum,
                                        coefficient_name = "took_hcqTRUE", transition = "1:3", adjusted = "age, sex, az, comorbidities (sum)"),
  frequentist_hypothesis_res_from_coxph(hypotheses$hcq_increases_discharged, fit_competing_hcq_az_comorbidities_sum,
                                        coefficient_name = "took_hcqTRUE", transition = "1:2", adjusted = "age, sex, az, comorbidities (sum)")
  )
hypothesis_res_list[["competing_hcq_az_comorbidities_sum"]]
```

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = paste0(cache_dir, "/competing_risks_res.csv"))
```

# Peak IL-6, peak D-dimer

# PCR positivity predicts

# Lymphocyte count predicts?