### 2020-09-20 manuscript/risk_scores.Rmd
Notes on risk scores

---
title: "Validating risk scores"
output: html_notebook
---

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)

devtools::load_all()
library(tidyverse)
library(patchwork)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- prepare_marker_data_wide(data)
```

# ACP

# Bello-Chavolla

- Requires pneumonia diagnosis (which we probably can't infer)... Maybe from using antibiotics?
- Early onset diabetes = diabetes & 40 years

- Age ≥ 65 years  3 
Diabetes  1
Diabetes*Age < 40 years 5
Age < 40 years -6
Obesity 1
Pneumonia 7
CKD 3
COPD 1
Immunosuppression 1

# Chen and Liu

𝑅𝑖𝑠𝑘𝑆𝑐𝑜𝑟𝑒𝑠=−4.211+0.013×𝐶𝑅𝑃+0.059×𝐴𝑔𝑒+0.112×𝐷𝐷−1.984×𝐿𝑌𝑀

Results for the validation dataset:
AUROC = 0.881
Cutoff (0.5)
Sensitivity = 0.839
Specificity = 0.794

```{r}

wide$ly
```

# Shi et al.

1 point for Age >= 50, Male, Hypertension

Not sure what "severe" means