### 2020-10-22 manuscript/analysis_brms.Rmd
Minor tweaks to brms analysis

---
title: "Analysis of Retrospective COVID-19 data"
author: "Paul Bürkner"
output:
  html_document:
    theme: default
encoding: UTF-8
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

---

# {.tabset}

## Setup

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(cache = TRUE, fig.width = 10)
```

```{r general_options, cache=FALSE}
devtools::load_all()
library(tidyverse)
library(tidybayes)
library(brms)
options(width = 160)
options(brms.backend = "cmdstanr")
options(mc.cores = 4)
theme_set(bayesplot::theme_default())

cache_dir <- here::here("local_temp_data", "analysis_brms")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
```

```{r functions}
scale2 <- function(x, na.rm = TRUE) {
  (x - mean(x, na.rm = na.rm)) / sd(x, na.rm = na.rm)
}

center <- function(x, na.rm = TRUE) {
  x - mean(x, na.rm = na.rm)
}

SW <- suppressWarnings

sum_coding <- function(x, lvls = levels(x)) {
  # codes the first category with -1
  nlvls <- length(lvls)
  stopifnot(nlvls > 1)
  cont <- diag(nlvls)[, -nlvls, drop = FALSE]
  cont[nlvls, ] <- -1
  cont <- cont[c(nlvls, 1:(nlvls - 1)), , drop = FALSE]
  colnames(cont) <- lvls[-1]
  x <- factor(x, levels = lvls)
  contrasts(x) <- cont
  x
}

dummy_coding <- function(x, lvls = levels(x)) {
  x <- factor(x, levels = lvls)
  contrasts(x) <- contr.treatment(levels(x))
  x
}

default_control <- list(adapt_delta = 0.95)
```

## Data preparation

```{r}
data <- read_data_for_analysis()
```

```{r}
data$patient_data <- data$patient_data %>%
  mutate(
    survival = outcome != "Death",
    discharged = outcome == "Discharged",
    outcome3 = fct_collapse(outcome, Discharged = c("Discharged", "Transferred")) %>%
      fct_relevel("Hospitalized", "Discharged", "Death")
  )
```

```{r}
data$surv_data <- data$marker_data_wide %>%  
  filter(day > 0) %>%
  mutate(
    daym1 = day - 1,
    status = fct_collapse(breathing, Hospitalized = breathing_levels) %>%
      factor(ordered = FALSE) %>% 
      fct_relevel("Hospitalized", "Discharged", "Death") %>%
      fct_explicit_na("Hospitalized"),
    censored = case_when(
      status == "Death" ~ "none",
      status %in% c("Hospitalized", "Discharged") ~ "right"
    ),
    censored_discharged = case_when(
      status == "Discharged" ~ "none",
      # approximation since patients who died cannot be 
      # discharged in the future
      status %in% c("Hospitalized", "Death") ~ "right"
    )
  ) %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id"))

data$surv_data_last_day <- data$surv_data %>%
  group_by(patient_id) %>%
  filter(day == max(day)) %>%
  ungroup()
```

```{r}
# we need to have "state at day 7", i.e. patients that died/were discharged 
# after day 7 are still treated as "Hospitalized" for the analysis. Other 
# studies (e.g. RECOVERY) work this way. Code provided by Martin.
surv_data_at_day <- function(patient_data, day) {
  res <- patient_data %>%
    filter(!(last_record < day & outcome3 == "Hospitalized"))

  # Non tidy to avoid hassle with factor levels  
  res$outcome3[res$last_record > day] <- "Hospitalized"
  res
}

data$surv_data_7 <- surv_data_at_day(data$surv_data, 7)
data$surv_data_28 <- surv_data_at_day(data$surv_data, 28)
```

```{r}
hypothesis_res_list <- list()
```

### Notes:

- days_from_symptom_onset. Days from symptom onset until hospitalization. 
Uncertain marker but contains some information. 
Required for anti viral (< 7 days) hypothesis.
- first_day_*: time from day 0 (arrived at hospital) to getting antivirals
- some sites only support patients with mild disease and some only with
strong disease symptoms

## Analysis: Survival

### Basic binary model

```{r, results="hide"}
form_survival_base <- bf(
  survival ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = bernoulli("logit")
)

prior_survival_base <- 
  prior(normal(0, 5), class = "b") +
  prior(normal(0, 5), class = "sd")

fit_survival_base <- brm(
  formula = form_survival_base,
  data = data$patient_data,
  prior = prior_survival_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_survival_base")
)
fit_survival_base <- add_criterion(fit_survival_base, "loo")
```

```{r}
summary(fit_survival_base)
# ce <- conditional_effects(fit_survival_base)
# plot(ce, ask = FALSE)
```

```{r}
#Marking the results as suspicious because there was a divergence...
draws <- tidy_draws(fit_survival_base)
hypothesis_res_list[["hcq_az_survival_base"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary",
    estimand = "logit(survival)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary",
    estimand = "logit(survival)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary",
    estimand = "logit(survival)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic binary model (7 days)

```{r, results="hide"}
fit_survival_base_7 <- brm(
  formula = form_survival_base,
  data = data$surv_data_7,
  prior = prior_survival_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_survival_base_7")
)
fit_survival_base_7 <- add_criterion(fit_survival_base_7, "loo")
```

```{r}
summary(fit_survival_base_7)
# ce <- conditional_effects(fit_survival_base_7)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_survival_base_7)
hypothesis_res_list[["hcq_az_survival_base_7"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary_7",
    estimand = "logit(survival)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary_7",
    estimand = "logit(survival)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary_7",
    estimand = "logit(survival)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic binary model (28 days)

```{r, results="hide"}
fit_survival_base_28 <- brm(
  formula = form_survival_base,
  data = data$surv_data_28,
  prior = prior_survival_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_survival_base_28")
)
fit_survival_base_28 <- add_criterion(fit_survival_base_28, "loo")
```

```{r}
summary(fit_survival_base_28)
# ce <- conditional_effects(fit_survival_base_28)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_survival_base_28)
hypothesis_res_list[["hcq_az_survival_base_28"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary_28",
    estimand = "logit(survival)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary_28",
    estimand = "logit(survival)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary_28",
    estimand = "logit(survival)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

## Analysis: Survival time

```{r, results="hide"}
form_survival_time <- bf(
  day | cens(censored) ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = brmsfamily("cox", "log")
)

prior_survival_time <- 
  prior(normal(0, 5), class = "b") +
  prior(normal(0, 5), class = "sd")

fit_survival_time <- brm(
  formula = form_survival_time,
  data = data$surv_data_last_day,
  prior = prior_survival_time,
  inits = 0,
  control = default_control,
  file = paste0(cache_dir, "/fit_survival_time")
)
fit_survival_time <- add_criterion(fit_survival_time, "loo")
```

```{r}
summary(fit_survival_time)
```

```{r}
draws <- tidy_draws(fit_survival_time)
hypothesis_res_list[["hcq_az_survival_time"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

## Analysis: Discharged

### Basic binary model

```{r, results="hide"}
form_discharged_base <- bf(
  discharged ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = bernoulli("logit")
)

prior_discharged_base <- 
  prior(normal(0, 5), class = "b") +
  prior(normal(0, 5), class = "sd")

fit_discharged_base <- brm(
  formula = form_discharged_base,
  data = data$patient_data,
  prior = prior_discharged_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_discharged_base")
)
fit_discharged_base <- add_criterion(fit_discharged_base, "loo")
```

```{r}
summary(fit_discharged_base)
# ce <- conditional_effects(fit_discharged_base)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_discharged_base)
hypothesis_res_list[["hcq_az_discharged_base"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic binary model (7 Days)

```{r, results="hide"}
fit_discharged_base_7 <- brm(
  formula = form_discharged_base,
  data = data$surv_data_7,
  prior = prior_discharged_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_discharged_base_7")
)
fit_discharged_base_7 <- add_criterion(fit_discharged_base_7, "loo")
```

```{r}
summary(fit_discharged_base_7)
# ce <- conditional_effects(fit_discharged_base_7)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_discharged_base_7)
hypothesis_res_list[["hcq_az_discharged_base_7"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic binary model (28 Days)

```{r, results="hide"}
fit_discharged_base_28 <- brm(
  formula = form_discharged_base,
  data = data$surv_data_28,
  prior = prior_discharged_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_discharged_base_28")
)
fit_discharged_base_28 <- add_criterion(fit_discharged_base_28, "loo")
```

```{r}
summary(fit_discharged_base_28)
# ce <- conditional_effects(fit_discharged_base_28)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_discharged_base_28)
hypothesis_res_list[["hcq_az_discharged_base_28"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_binary_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_binary_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiravirTRUE,
    model = "brms_binary_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

## Analysis: discharged time

```{r, results="hide"}
form_discharged_time <- bf(
  day | cens(censored_discharged) ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = brmsfamily("cox", "log")
)

prior_discharged_time <- 
  prior(normal(0, 5), class = "b") +
  prior(normal(0, 5), class = "sd")

fit_discharged_time <- brm(
  formula = form_discharged_time,
  data = data$surv_data_last_day,
  prior = prior_discharged_time,
  inits = 0,
  control = list(adapt_delta = 0.98),
  file = paste0(cache_dir, "/fit_discharged_time")
)
fit_discharged_time <- add_criterion(fit_discharged_time, "loo")
```

```{r}
summary(fit_discharged_time)
```

```{r}
draws <- tidy_draws(fit_discharged_time)
hypothesis_res_list[["hcq_az_discharged_time"]] <- rbind(
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_hcqTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_azTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_ever_favipiraviTRUE,
    model = "brms_cox",
    estimand = "log(HR)",
    hypothesis = hypotheses$favipiravi_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

## Analysis: Hospitalized vs. Discharged vs. Death

### Basic Categorical model

```{r, results="hide"}
form_outcome_base <- bf(
  outcome3 ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = categorical("logit")
)

prior_outcome_base <- 
  prior(normal(0, 5), class = "b", dpar = "muDeath") +
  prior(normal(0, 5), class = "sd", dpar = "muDeath") +
  prior(normal(0, 5), class = "b", dpar = "muDischarged") +
  prior(normal(0, 5), class = "sd", dpar = "muDischarged")

fit_outcome_base <- brm(
  formula = form_outcome_base,
  data = data$patient_data,
  prior = prior_outcome_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_outcome_base")
)
fit_outcome_base <- add_criterion(fit_outcome_base, "loo")
```

```{r}
summary(fit_outcome_base)
# ce <- conditional_effects(fit_outcome_base)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_outcome_base)
hypothesis_res_list[["hcq_az_outcome_base"]] <- rbind(
  # Death 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_hcqTRUE,
    model = "brms_categorical",
    estimand = "logit(death)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_azTRUE,
    model = "brms_categorical",
    estimand = "logit(death)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_favipiravirTRUE,
    model = "brms_categorical",
    estimand = "logit(death)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  ),
  # Discharged
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_hcqTRUE,
    model = "brms_categorical",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_azTRUE,
    model = "brms_categorical",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_favipiravirTRUE,
    model = "brms_categorical",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic Categorical model (7 Days)

```{r, results="hide"}
fit_outcome_base_7 <- brm(
  formula = form_outcome_base,
  data = data$surv_data_7,
  prior = prior_outcome_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_outcome_base_7")
)
fit_outcome_base_7 <- add_criterion(fit_outcome_base_7, "loo")
```

```{r}
summary(fit_outcome_base_7)
# ce <- conditional_effects(fit_outcome_base_7)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_outcome_base_7)
hypothesis_res_list[["hcq_az_outcome_base_7"]] <- rbind(
  # Death 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_hcqTRUE,
    model = "brms_categorical_7",
    estimand = "logit(death)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_azTRUE,
    model = "brms_categorical_7",
    estimand = "logit(death)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_favipiravirTRUE,
    model = "brms_categorical_7",
    estimand = "logit(death)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  ),
  # Discharged
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_hcqTRUE,
    model = "brms_categorical_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_azTRUE,
    model = "brms_categorical_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_favipiravirTRUE,
    model = "brms_categorical_7",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

### Basic Categorical model (28 Days)

```{r, results="hide"}
fit_outcome_base_28 <- brm(
  formula = form_outcome_base,
  data = data$surv_data_28,
  prior = prior_outcome_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_outcome_base_28")
)
fit_outcome_bas_28 <- add_criterion(fit_outcome_base_28, "loo")
```

```{r}
summary(fit_outcome_base_28)
# ce <- conditional_effects(fit_outcome_base_28)
# plot(ce, ask = FALSE)
```

```{r}
draws <- tidy_draws(fit_outcome_base_28)
hypothesis_res_list[["hcq_az_outcome_base_28"]] <- rbind(
  # Death
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_hcqTRUE,
    model = "brms_categorical_28",
    estimand = "logit(death)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_azTRUE,
    model = "brms_categorical_28",
    estimand = "logit(death)",
    hypothesis = hypotheses$az_reduces_death,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDeath_ever_favipiravirTRUE,
    model = "brms_categorical_28",
    estimand = "logit(death)",
    hypothesis = hypotheses$favipiravir_reduces_death,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  ),
  # Discharged
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_hcqTRUE,
    model = "brms_categorical_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "az, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_azTRUE,
    model = "brms_categorical_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$az_increases_discharged,
    adjusted = "hcq, favipiravir, sex, age_norm, smoking, obesity"
  ),
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_muDischarged_ever_favipiravirTRUE,
    model = "brms_categorical_28",
    estimand = "logit(discharged)",
    hypothesis = hypotheses$favipiravir_increases_discharged,
    adjusted = "hcq, az, sex, age_norm, smoking, obesity"
  )
)
```

## Analysis: Worst Breathing {.tabset}

### Basic ordinal model

```{r, results="hide"}
form_worst_breathing_base <- bf(
  worst_breathing ~ sex + age_norm + 
    smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir +
    (1 | hospital_id),
  family = sratio("logit")
)

prior_worst_breathing_base <- 
  prior(normal(0, 5), class = "b") +
  prior(normal(0, 5), class = "sd")

fit_worst_breathing_base <- brm(
  formula = form_worst_breathing_base,
  data = data$patient_data,
  prior = prior_worst_breathing_base,
  control = default_control,
  file = paste0(cache_dir, "/fit_worst_breathing_base")
)
fit_worst_breathing_base <- add_criterion(fit_worst_breathing_base, "loo")
```

```{r}
summary(fit_worst_breathing_base)
# ce <- conditional_effects(fit_worst_breathing_base, categorical = TRUE)
# plot(ce, ask = FALSE)
```

### Category-specific ordinal model

```{r, results="hide"}
# form_worst_breathing_cs <- bf(
#   worst_breathing ~ cs(sex + age_norm + 
#     smoking + mo(obesity) + ever_hcq + ever_az + ever_favipiravir) +
#     (1 | hospital_id),
#   family = sratio("logit")
# )
# 
# prior_worst_breathing_cs <- 
#   prior(normal(0, 5), class = "b") +
#   prior(normal(0, 5), class = "sd")
# 
# fit_worst_breathing_cs <- brm(
#   formula = form_worst_breathing_cs,
#   data = data$patient_data,
#   prior = prior_worst_breathing_cs,
#   inits = 0,
#   file = paste0(cache_dir, "/fit_worst_breathing_cs")
# )
# fit_worst_breathing_cs <- add_criterion(fit_worst_breathing_cs, "loo")
```

```{r}
# summary(fit_worst_breathing_cs)
# ce <- conditional_effects(fit_worst_breathing_cs, categorical = TRUE)
# plot(ce, ask = FALSE)
```

### Model Comparison

```{r}
# loo(fit_worst_breathing_base, fit_worst_breathing_cs)
```

## Store Hypotheses

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = here::here("manuscript/analysis_brms.csv"))
```