### 2020-10-27 manuscript/single_outcome_coxph.Rmd
TS's stuff added: treatment combinations overview, baseline associations with treatment, convenience sample warning, coxph modeling

---
title: "Survival - frequentist, single outcome"
output: html_notebook
---

```{r}
devtools::load_all()
library(tidyverse)
library(survival)

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()

# add hcq_first
pd<-data$patient_data
pd$hcq_first<-NA
for (i in unique(pd$patient_id)) {
  tmp<-data$marker_data$day[data$marker_data$patient_id==i&data$marker_data$marker=='hcq']
  if (sum(!is.na(tmp))>0) {
    pd$hcq_first[pd$patient_id==i]<-min(tmp)
  }
}

# prepare surv data
d<-c()
# I know, the code is busy, but intuitive
for (i in unique(pd$patient_id)) {
  cnd<-pd$patient_id==i
  if (pd$last_record[cnd]==0) next
  hcq_first<-pd$hcq_first[cnd]
  if (!is.na(hcq_first)) {
    if (hcq_first>0) {
      if (pd$last_record[cnd]>hcq_first) {
        if (hcq_first>1) {
          d<-rbind(d,data.frame(t0=0,t1=hcq_first-1,outcome=0,patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],hcq=0))
        }
        d<-rbind(d,data.frame(t0=hcq_first-1,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],hcq=1))
      } else {
        d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],hcq=0))
      }
    } else { # hcq_first=0
      d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],hcq=1))
    }
  } else {
    d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],hcq=0))
  }
}
d<-within(d,comorb_sum<-ischemic_heart_disease+has_hypertension_drugs+heart_failure+COPD+    other_lung_disease+    renal_disease)

hypothesis_res_list <- list()

cache_dir <- here::here("local_temp_data")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}

```

# Risk of death, increased discharge

```{r}

m1<-with(d,coxph(S<-Surv(t0,t1,outcome=='Death')~hcq+age+sex+strata(hospital_id)+comorb_sum))
m0<-with(d,coxph(S<-Surv(t0,t1,outcome=='Death')~    age+sex+strata(hospital_id)+comorb_sum))

mt1<-with(d,coxph(S<-Surv(t0,t1,outcome=='Discharged')~hcq+age+sex+strata(hospital_id)+comorb_sum))
mt0<-with(d,coxph(S<-Surv(t0,t1,outcome=='Discharged')~    age+sex+strata(hospital_id)+comorb_sum))

hypothesis_res_list[["competing_hcq_comorbidities_sum"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_reduces_death, 
                                         point_estimate = coef(m1)[1],
                                         adjusted = "age, sex, comorbidities (sum)", 
                                         test_stat = anova(m1,m0,test='Chisq')[2,2],
                                         df = anova(m1,m0,test='Chisq')[2,3],
                                         p = anova(m1,m0,test='Chisq')[2,4],
                                         ci_low = confint(m1)[1,1],
                                         ci_high = confint(m1)[1,2]),
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_increases_discharged, 
                                         point_estimate = coef(mt1)[1],
                                         adjusted = "age, sex, comorbidities (sum)", 
                                         test_stat = anova(mt1,mt0,test='Chisq')[2,2],
                                         df = anova(mt1,mt0,test='Chisq')[2,3],
                                         p = anova(mt1,mt0,test='Chisq')[2,4],
                                         ci_low = confint(mt1)[1,1],
                                         ci_high = confint(mt1)[1,2])
  )
hypothesis_res_list[["competing_hcq_comorbidities_sum"]]

```

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = here::here("manuscript", "/single_outcome_coxph_res.csv"))
```