### 2020-10-31 manuscript/hmm.Rmd
Updated HMM results

---
title: "HMM Model"
output: html_document
---

```{r setup}
devtools::load_all()
library(tidyverse)
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)
options(brms.backend = "cmdstanr")

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()

if(length(unique(data$patient_data$patient_id)) != length(data$patient_data$patient_id)) {
  stop("FAil")
}
# warning("Subsetting data")
# set.seed(4568224)
# patients_to_use <- sample(data$patient_data$patient_id, size = 70)
# data <- filter_complete_data(data, patient_id %in% patients_to_use)

wide <- data$marker_data_wide

cache_dir <- here::here("local_temp_data", "hmm")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

do_pp_checks <- FALSE
hmm_hypothesis_res_list <- list()

cl <-  parallel::makePSOCKcluster(parallel::detectCores())

#Temporary hack
registerS3method("stan_log_lik", class = "rate_hmm", method = stan_log_lik.rate_hmm, envir = asNamespace("brms"))
```

```{r}
transform_serie_data <- function(raw_serie_data) {
  raw_serie_data  %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care),
         is_male = as.integer(sex == "M")
         )
}

serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id"), suffix = c("_admission", "")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

serie_data_28 <- wide %>%
  right_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day"),  suffix = c("_admission", "")) %>%
  # group_by(patient_id) %>%
  # mutate(
  #   across(
  #     c(starts_with("took_"), all_of("best_supportive_care")),
  #     ~ if_else(is.na(.x) & day > max(day[!is.na(.x)]), .x[which.max(day[!is.na(.x)])], .x)
  #     ),
  # ) %>%
  # ungroup() %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

# Extend terminal states and took_XX data for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  mutate(across(c(tidyselect::starts_with("took_"), all_of("best_supportive_care")), ~ if_else(is.na(.x), as.integer(any(.x != 0, na.rm = TRUE)), .x))) %>%
  ungroup()

# Check succesful extension
mismatches <- serie_data_28 %>% inner_join(serie_data, by = c("patient_id", "day")) %>%
  select(-starts_with("."), -last_day,  -last_state) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(!any_of(c("patient_id", "day")), names_to = c("col", "source"), names_sep = "\\.") %>%
  group_by(patient_id, day, col) %>%
  summarise(n_vals = length(unique(value)), .groups = "drop") %>%
  filter(n_vals > 1)

if(nrow(mismatches) > 0) {
  print(mismatches)
  stop("Mismatches in serie_data_28")
}

base_observed_state_data <- tibble(id = factor(disease_s_levels, levels = disease_s_levels), is_noisy = FALSE)

simple_hidden_state_data <- tibble(id = factor(disease_s_levels, levels = disease_s_levels)) %>% mutate( corresponding_obs = id)

simple_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

default_control <- list(adapt_delta = 0.9)
```

```{r}
#data$patient_data %>% filter(!is.na(best_supportive_care_from)) %>% select(outcome)
```

```{r}
recode_group <- function(rate_data, group_name) {
  is_column <- paste0("is_", group_name) 
  id_column <- paste0(group_name, "_id")

  rate_data %>% mutate(
    !!is_column := as.double(rate_group == group_name),
    !!id_column := factor(
      if_else(rate_group == group_name, as.integer(.rate_id) - min(as.integer(.rate_id[rate_group == group_name])), as.integer(0)), 
      levels = 0:(sum(rate_group == group_name) - 1), 
      ordered = TRUE)
  )
}

```

# Treatments - effects on rate groups

## Only treatments

```{r trt_group_m1}
trt_group_m1_hidden_state_data <- simple_hidden_state_data

trt_group_m1_observed_state_data <- base_observed_state_data

simple_rate_data <- rbind(
  tibble(.from = c("Oxygen", "Ventilated"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[2 : length(breathing_s_levels)], .to = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], .to = breathing_s_levels[2 : length(breathing_s_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
         .from = factor(.from, levels = disease_s_levels),
         .to = factor(.to, levels = disease_s_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"),
         )

simple_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

trt_group_m1_prior = #brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(-2, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

trt_group_m1_formula <- ~ 0 + .rate_id +  (0 + best_supportive_care + took_hcq + took_az + took_favipiravir + took_convalescent_plasma || rate_group)

trt_group_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  simple_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_group_m1_fit <- brmhmm(trt_group_m1, cache_file = paste0(cache_dir,"/trt_group_m1.rds"), control = default_control, init = 0.1, iter = 1500)

summary(trt_group_m1_fit$brmsfit)
```

```{r, fig.width=12, fig.height=12}
#bayesplot::mcmc_pairs(xx$fit, regex_pars = c("b_.rate_id"))
# bayesplot::mcmc_pairs(trt_group_m1_fit$brmsfit, regex_pars = "worsen_one", np = nuts_params(trt_group_m2_fit$brmsfit))

```

```{r, fig.width=12, fig.height=12}
# bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, pars = c("b_.rate_idAA_Discharged", "b_.rate_idOxygen_AA"), regex_pars = "improve.*(age|male)")
# 
#bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, regex_pars = c("b_.rate_id"))
```

```{r}
# fit <- trt_group_m1_fit
# data_28 <- trt_group_m1
# data_28$serie_data <- serie_data_28
# epred_rect <- posterior_epred_rect(fit)#, newdata = data_28)
# predicted_rect <- posterior_epred_to_predicted(fit, epred_rect)
# # predicted <- posterior_rect_to_long(fit, predicted_rect)
# 
# pp_check_state_at(fit, predicted_rect, 28, newdata = data_28)
```

```{r}

```

```{r}
hmm_hypothesis_res_list[["trt_group_m1"]] <- evaluate_all_treatment_hypotheses(trt_group_m1_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive", model_check = "Suspicious", cl = cl)
```

```{r}
read_csv("D:/Google Drive/MBÚ/Covid/hmm_res.csv")
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

## Treatments + age + sex

```{r trt_group_m2}
trt_group_m2_formula <- update.formula(trt_group_m1_formula,  ~ . +  (0 + age_norm + is_male || rate_group))
trt_group_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m2_fit <- brmhmm(trt_group_m2, cache_file = paste0(cache_dir,"/trt_group_m2.rds"), init = 0.1, control = default_control, iter = 1200)

#parnames(trt_group_m2_fit$brmsfit)
summary(trt_group_m2_fit$brmsfit)
```
```{r, fig.width=12, fig.height=12}
# 
#  
# bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, regex_pars = "improve_one", np = nuts_params(trt_group_m2_fit$brmsfit))
# 
# 
# bayesplot::mcmc_parcoord(trt_group_m2_fit$brmsfit, regex_pars = "death", np = nuts_params(trt_group_m2_fit$brmsfit))
# bayesplot::mcmc_parcoord(trt_group_m2_fit$brmsfit, regex_pars = "worsen_one", np = nuts_params(trt_group_m2_fit$brmsfit))
```

```{r}

hmm_hypothesis_res_list[["trt_group_m2"]] <- evaluate_all_treatment_hypotheses(trt_group_m2_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m2_fit)
}
```

## Treatments + age + sex + hospital

```{r trt_group_m3}
trt_group_m3_formula <- update.formula(trt_group_m2_formula,  ~  . + (0 + .rate_id | hospital_id))

trt_group_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m3_fit <- brmhmm(trt_group_m3, cache_file = paste0(cache_dir,"/trt_group_m3.rds"), init = 0.1, control = default_control, iter = 1200)

summary(trt_group_m3_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_group_m3"]] <-  evaluate_all_treatment_hypotheses(trt_group_m3_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "OK", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m3_fit)
}
```

## Treatment + age + sex + hospital + comorbidities

```{r trt_group_m4}
trt_group_m4_formula <-  update.formula(trt_group_m3_formula,  ~ . + (comorbidities_sum || rate_group))

trt_group_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m4_fit <- brmhmm(trt_group_m4, cache_file = paste0(cache_dir,"/trt_group_m4.rds"), init = 0.1, control = default_control, iter = 1200)

summary(trt_group_m4_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_group_m4"]] <-  evaluate_all_treatment_hypotheses(trt_group_m4_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "OK", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m4_fit)
}
```

# Treatments - effect on individual rates

## Only treatments

```{r trt_rate_m1}
trt_rate_m1_hidden_state_data <- simple_hidden_state_data

trt_rate_m1_observed_state_data <- base_observed_state_data

trt_rate_m1_rate_data <- simple_rate_data

trt_rate_m1_initial_states <- simple_initial_states

trt_rate_m1_prior = trt_group_m1_prior

trt_rate_m1_formula <- ~ 0 + .rate_id +  (0 + best_supportive_care + took_hcq + took_az + took_favipiravir + took_convalescent_plasma || .rate_id)

trt_rate_m1 <- brmshmmdata(trt_rate_m1_formula, serie_data, 
                  trt_rate_m1_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_rate_m1_fit <- brmhmm(trt_rate_m1, cache_file = paste0(cache_dir,"/trt_rate_m1.rds"), control = default_control, init = 0.1)

summary(trt_rate_m1_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_rate_m1"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m1_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive", model_check = "Suspicious", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_rate_m1_fit)
}
```

## Treatments + age + sex

```{r trt_rate_m2}
trt_rate_m2_formula <- update.formula(trt_rate_m1_formula,  ~ . +  (0 + age_norm + is_male || .rate_id))
trt_rate_m2 <- brmshmmdata(trt_rate_m2_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m2_fit <- brmhmm(trt_rate_m2, cache_file = paste0(cache_dir,"/trt_rate_m2.rds"), init = 0.1, control = default_control)

summary(trt_rate_m2_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_rate_m2"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m2_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_rate_m2_fit)
}
```

## Treatments + age + sex + hospital

```{r trt_rate_m3}
trt_rate_m3_formula <- update.formula(trt_rate_m2_formula,  ~  . + (0 + .rate_id | hospital_id))
trt_rate_m3 <- brmshmmdata(trt_rate_m3_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m3_fit <- brmhmm(trt_rate_m3, cache_file = paste0(cache_dir,"/trt_rate_m3.rds"), init = 0.1, control = default_control, iter = 1500)

summary(trt_rate_m3_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_rate_m3"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m3_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "OK", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_rate_m3_fit)
}
```

## Treatment + age + sex + hospital + comorbidities

```{r trt_rate_m4}
trt_rate_m4_formula <-  update.formula(trt_rate_m3_formula, ~ . + (comorbidities_sum || .rate_id))

trt_rate_m4 <- brmshmmdata(trt_rate_m4_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m4_fit <- brmhmm(trt_rate_m4, cache_file = paste0(cache_dir,"/trt_rate_m4.rds"), init = 0.1, control = default_control, iter = 1500)

summary(trt_rate_m4_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["trt_rate_m4"]] <- evaluate_all_treatment_hypotheses(trt_rate_m4_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "OK", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_rate_m4_fit)
}
```

# Markers

## D-dimer + hospital

```{r d_dimer_m1}
d_dimer_serie_data <- serie_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

d_dimer_initial_states <- d_dimer_serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

d_dimer_m1_formula <-  ~ 0 + .rate_id +  (0 + best_supportive_care || rate_group)  + (0 + .rate_id | hospital_id) + (log_peak_d_dimer || rate_group)

d_dimer_m1 <- brmshmmdata(d_dimer_m1_formula, d_dimer_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m1_fit <- brmhmm(d_dimer_m1, cache_file = paste0(cache_dir,"/d_dimer_m1.rds"), init = 0.1, control = default_control, iter = 1500)

summary(d_dimer_m1_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(d_dimer_m1_fit$brmsfit)
hmm_hypothesis_res_list[["d_dimer_m1_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$`r_rate_group[death,log_peak_d_dimer]`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$d_dimer_increases_death,
    adjusted = "supportive, hospital",
    model_check = "Suspicious"
  )

```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(d_dimer_m1_fit)
}
```

## D-dimer + age + sex + comorbidities

```{r}
d_dimer_m2_formula <-  update.formula(d_dimer_m1_formula, ~ . + (0 + age_norm + is_male || rate_group))

d_dimer_m2 <- brmshmmdata(d_dimer_m2_formula, d_dimer_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m2_fit <- brmhmm(d_dimer_m2, cache_file = paste0(cache_dir,"/d_dimer_m2.rds"), init = 0.1, control = default_control, iter = 1500)

summary(d_dimer_m2_fit$brmsfit)

```

```{r}
draws <- tidybayes::tidy_draws(d_dimer_m2_fit$brmsfit)
hmm_hypothesis_res_list[["d_dimer_m2_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$`r_rate_group[death,log_peak_d_dimer]`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$d_dimer_increases_death,
    adjusted = "supportive, age, sex, hospital",
    model_check = "Suspicious"
  )

```

## IL-6 + hospital

```{r}
IL_6_serie_data <- serie_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "IL_6", new_column_name = "peak_IL_6", initial_value = 1) %>%
  mutate(log_peak_IL_6 = log(peak_IL_6))

IL_6_initial_states <- IL_6_serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

IL_6_m1_formula <- ~ 0 + .rate_id +  (0 + best_supportive_care || rate_group)  + (0 + .rate_id | hospital_id) + (log_peak_IL_6 || rate_group)

IL_6_m1 <- brmshmmdata(IL_6_m1_formula, IL_6_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m1_fit <- brmhmm(IL_6_m1, cache_file = paste0(cache_dir,"/IL_6_m1.rds"), init = 0.1, control = default_control, iter = 1500)

summary(IL_6_m1_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(IL_6_m1_fit$brmsfit)
hmm_hypothesis_res_list[["IL_6_m1_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$`r_rate_group[death,log_peak_IL_6]`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$IL_6_increases_death,
    adjusted = "supportive, hospital"
  )

```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(IL_6_m1_fit)
}
```

## IL_6 + age + sex + comorbidities

```{r}
IL_6_m2_formula <-  update.formula(IL_6_m1_formula, ~ . + (0 + age_norm + is_male || rate_group))

IL_6_m2 <- brmshmmdata(IL_6_m2_formula, IL_6_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m2_fit <- brmhmm(IL_6_m2, cache_file = paste0(cache_dir,"/IL_6_m2.rds"), init = 0.1, control = default_control, iter = 1500)

summary(IL_6_m2_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(IL_6_m2_fit$brmsfit)
hmm_hypothesis_res_list[["IL_6_m2_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$`r_rate_group[death,log_peak_IL_6]`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$IL_6_increases_death,
    adjusted = "supportive, age, sex, hospital"
  )

```

# Complex: Treatments - effects on rate groups

## Only treatments

```{r c_trt_m1}
complex_rate_data <- rbind(
  tibble(.from = paste0(c("Oxygen", "Ventilated"), "_worsening"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA_improving", 
         .to = "Discharged", 
         rate_group = "improve_one"),
  tibble(.from = paste0(breathing_s_levels[2 : length(breathing_s_levels)], "_improving"), 
         .to = paste0(breathing_s_levels[1 : (length(breathing_s_levels) - 1)] , "_improving"), 
         rate_group = "improve_one"),
  tibble(.from = paste0(breathing_s_levels[1 : (length(breathing_s_levels) - 1)], "_worsening"),
         .to = paste0(breathing_s_levels[2 : length(breathing_s_levels)], "_worsening"),
         rate_group = "worsen_one"),
  tibble(.from = paste0(breathing_s_levels, "_worsening"),
         .to = paste0(breathing_s_levels, "_improving"),
         rate_group = "to_improving"),
  tibble(.from = paste0(breathing_s_levels, "_improving"),
         .to = paste0(breathing_s_levels, "_worsening"),
         rate_group = "to_worsening")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
  ) %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"))

complex_hidden_state_data <- rbind(
  tibble(base = breathing_s_levels) %>% mutate(corresponding_obs = base) %>% 
    crossing(tibble(state_group = c("_improving", "_worsening"))) %>% 
    mutate(id = factor(paste0(base, state_group))) %>%
    select(-base, -state_group),
  tibble(id = c("Death","Discharged")) %>% mutate(corresponding_obs = id)
)

complex_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")

c_trt_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m1_fit <- brmhmm(c_trt_m1, cache_file = paste0(cache_dir,"/c_trt_m1.rds"), control = default_control, init = 0.1, iter = 1500)

summary(c_trt_m1_fit$brmsfit)

```

```{r}
hmm_hypothesis_res_list[["c_trt_rate_m1"]] <- evaluate_all_treatment_hypotheses(c_trt_m1_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive", model_check = "Suspicious", cl = cl)
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m1_fit)
}
```

## Treatments + age + sex

```{r c_trt_m2}
c_trt_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m2_fit <- brmhmm(c_trt_m2, cache_file = paste0(cache_dir,"/c_trt_m2.rds"), init = 0.1, control = default_control, iter = 1200)

#parnames(c_trt_m2_fit$brmsfit)
summary(c_trt_m2_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["c_trt_rate_m2"]] <- evaluate_all_treatment_hypotheses(c_trt_m2_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious", cl = cl)

```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m2_fit)
}
```

## Treatments + age + sex + hospital

```{r c_trt_m3}
c_trt_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m3_fit <- brmhmm(c_trt_m3, cache_file = paste0(cache_dir,"/c_trt_m3.rds"), init = 0.1, control = default_control, iter = 1200)

summary(c_trt_m3_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["c_trt_rate_m3"]] <- evaluate_all_treatment_hypotheses(c_trt_m3_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "Suspicious", cl = cl)

```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m3_fit)
}
```

## Treatment + age + sex + hospital + comorbidities

```{r c_trt_m4}
c_trt_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m4_fit <- brmhmm(c_trt_m4, cache_file = paste0(cache_dir,"/c_trt_m4.rds"), init = 0.1, control = default_control, iter = 1200)

summary(c_trt_m4_fit$brmsfit)
```

```{r}
hmm_hypothesis_res_list[["c_trt_rate_m4"]] <- evaluate_all_treatment_hypotheses(c_trt_m4_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "Suspicious", cl = cl)

```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m4_fit)
}
```

# Environment

```{r}
hmm_hypothesis_res_all <- do.call(rbind, hmm_hypothesis_res_list)
write_csv(hmm_hypothesis_res_all, path = here::here("manuscript", "hmm_res.csv"))
```

```{r}
sessionInfo()
```