### 2020-11-21 manuscript/supplement.Rmd

---
site: "bookdown::bookdown_site"
title: "Supplementary material for Detailed disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis"
output:
  bookdown::pdf_document2
bibliography: "Covid19.json"
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
#author: "`r { source(here::here('manuscript', 'authors.R')); get_authors_string() }`"
---

```{r setup, echo=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache = TRUE, cache.lazy = FALSE)
```

\newpage

```{r child_hmm, child = 'hmm.Rmd'}
```