### 2020-11-26 manuscript/single_outcome_coxph.Rmd

---
title: "Survival - frequentist, single outcome"
output:
  bookdown::pdf_document2
---

# Single outcome proprtional hazards model

Here we use a single outcome survival model via the `coxph` function of the `Survival` package. Hydroxychloroquine is treated as a time-varying covariate, other covariates are fixed.

```{r setup_coxph_single, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
devtools::load_all(here::here())
library(tidyverse)
library(survival)

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()

# add hcq_first
pd<-data$patient_data
pd$hcq_first<-NA
for (i in unique(pd$patient_id)) {
  tmp<-data$marker_data$day[data$marker_data$patient_id==i&data$marker_data$marker=='hcq']
  if (sum(!is.na(tmp))>0) {
    pd$hcq_first[pd$patient_id==i]<-min(tmp)
  }
}

# prepare surv data
d<-c()
# I know, the code is busy, but intuitive
for (i in unique(pd$patient_id)) {
  cnd<-pd$patient_id==i
  if (pd$last_record[cnd]==0) next
  hcq_first<-pd$hcq_first[cnd]
  if (!is.na(hcq_first)) {
    if (hcq_first>0) {
      if (pd$last_record[cnd]>hcq_first) {
        if (hcq_first>1) {
          d<-rbind(d,data.frame(t0=0,t1=hcq_first-1,outcome=0,patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],first_wave=pd$first_wave[cnd],hcq=0))
        }
        d<-rbind(d,data.frame(t0=hcq_first-1,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],first_wave=pd$first_wave[cnd],hcq=1))
      } else {
        d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],first_wave=pd$first_wave[cnd],hcq=0))
      }
    } else { # hcq_first=0
      d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],first_wave=pd$first_wave[cnd],hcq=1))
    }
  } else {
    d<-rbind(d,data.frame(t0=0,t1=pd$last_record[cnd],outcome=pd$outcome[cnd],patientId=i,hcq_first=pd$hcq_first[cnd],last_record=pd$last_record[cnd],age=pd$age[cnd],sex=pd$sex[cnd],hospital_id=pd$hospital_id[cnd],admitted_for_covid=pd$admitted_for_covid[cnd],ischemic_heart_disease=pd$ischemic_heart_disease[cnd],has_hypertension_drugs=pd$has_hypertension_drugs[cnd],heart_failure=pd$heart_failure[cnd],renal_disease=pd$renal_disease[cnd],high_creatinin=pd$high_creatinin[cnd],COPD=pd$COPD[cnd],other_lung_disease=pd$other_lung_disease[cnd],renal_disease=pd$renal_disease[cnd],obesity=pd$obesity[cnd],first_wave=pd$first_wave[cnd],hcq=0))
  }
}
d<-within(d,comorb_sum<-ischemic_heart_disease+has_hypertension_drugs+heart_failure+COPD+    other_lung_disease+    renal_disease)

hypothesis_res_list <- list()

```

## Risk of death, increased discharge

Here, we compare the model with hydroxychloroquine as a covariate (`m1`) with model not including hydroxychloroquine (`m0`), the formulas used are:

```
m1: Surv(t0,t1,outcome=='Death')~hcq+age+sex+strata(hospital_id)+comorb_sum)
m0: Surv(t0,t1,outcome=='Death')~    age+sex+strata(hospital_id)+comorb_sum)
```

where `comorb_sum` is a score representing the cumulative presence of ischemic heart disease, hypertension drugs, former heart failure, COPD, other lung diseases, renal disease, or high creatinine. A separate model is fitted for the "Discharged" outcome.

The point estimate of the effect and its confidence intervals are taken directly from the `m1` model, a p-value is computed via a call `anova` comparing the two models.

Here are the fitted models for the "Death" outcome:

```{r}

m1<-with(d,coxph(S<-Surv(t0,t1,outcome=='Death')~hcq+age+sex+strata(hospital_id)+comorb_sum))
m0<-with(d,coxph(S<-Surv(t0,t1,outcome=='Death')~    age+sex+strata(hospital_id)+comorb_sum))

m1
m0

```

And the fitted models for the "Discharged" outcome:

```{r}
mt1<-with(d,coxph(S<-Surv(t0,t1,outcome=='Discharged')~hcq+age+sex+strata(hospital_id)+comorb_sum))
mt0<-with(d,coxph(S<-Surv(t0,t1,outcome=='Discharged')~    age+sex+strata(hospital_id)+comorb_sum))

mt1
mt0
```

And here are the resulting evaluation of individual hypotheses:

```{r}
hypothesis_res_list[["competing_hcq_comorbidities_sum"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_death, 
                                         point_estimate = coef(m1)[1],
                                         adjusted = "age, sex, hospital, comorbidities (sum)", 
                                         test_stat = anova(m1,m0,test='Chisq')[2,2],
                                         df = anova(m1,m0,test='Chisq')[2,3],
                                         p = anova(m1,m0,test='Chisq')[2,4],
                                         ci_low = confint(m1)[1,1],
                                         ci_high = confint(m1)[1,2]),
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_hospital, 
                                         point_estimate = -coef(mt1)[1],
                                         adjusted = "age, sex, hospital, comorbidities (sum)", 
                                         test_stat = anova(mt1,mt0,test='Chisq')[2,2],
                                         df = anova(mt1,mt0,test='Chisq')[2,3],
                                         p = anova(mt1,mt0,test='Chisq')[2,4],
                                         ci_low = -confint(mt1)[1,2],
                                         ci_high = -confint(mt1)[1,1])
  )
hypothesis_res_list[["competing_hcq_comorbidities_sum"]] %>% select(hypothesis, point_estimate, p_value, ci_low, ci_high)

```

## Risk of death, increased discharge - first wave only

Here, we perform the same analysis as above, but including only patients from the first wave, here are the fitted models for the "Death" outcome:

```{r}
d_first <- d %>% filter(first_wave)
m1_first<-with(d_first,coxph(S<-Surv(t0,t1,outcome=='Death')~hcq+age+sex+strata(hospital_id)+comorb_sum))
m0_first<-with(d_first,coxph(S<-Surv(t0,t1,outcome=='Death')~    age+sex+strata(hospital_id)+comorb_sum))

m1_first
m0_first

```

And here are the fits for the "Discharged" outcome:

```{r}
mt1_first<-with(d_first,coxph(S<-Surv(t0,t1,outcome=='Discharged')~hcq+age+sex+strata(hospital_id)+comorb_sum))
mt0_first<-with(d_first,coxph(S<-Surv(t0,t1,outcome=='Discharged')~    age+sex+strata(hospital_id)+comorb_sum))

mt1_first
mt0_first
```

Which give us the following evaluation of the hypotheses:

```{r}
hypothesis_res_list[["competing_hcq_comorbidities_sum_fw"]] <-
  rbind(
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_death, 
                                         point_estimate = coef(m1_first)[1],
                                         adjusted = "age, sex, hospital, comorbidities (sum), first_wave", 
                                         test_stat = anova(m1_first,m0_first,test='Chisq')[2,2],
                                         df = anova(m1_first,m0_first,test='Chisq')[2,3],
                                         p = anova(m1_first,m0_first,test='Chisq')[2,4],
                                         ci_low = confint(m1_first)[1,1],
                                         ci_high = confint(m1_first)[1,2]),
  frequentist_hypothesis_res_from_coxph1(hypotheses$hcq_hospital, 
                                         point_estimate = -coef(mt1_first)[1],
                                         adjusted = "age, sex, hospital, comorbidities (sum), first_wave", 
                                         test_stat = anova(mt1_first,mt0_first,test='Chisq')[2,2],
                                         df = anova(mt1_first,mt0_first,test='Chisq')[2,3],
                                         p = anova(mt1_first,mt0_first,test='Chisq')[2,4],
                                         ci_low = -confint(mt1_first)[1,2],
                                         ci_high = -confint(mt1_first)[1,1])
  )

hypothesis_res_list[["competing_hcq_comorbidities_sum_fw"]] %>% select(hypothesis, point_estimate, p_value, ci_low, ci_high)

```

```{r}
hypothesis_res_all <- do.call(rbind, hypothesis_res_list)
write_csv(hypothesis_res_all, path = here::here("manuscript", "/single_outcome_coxph_res.csv"))
```