### 2021-01-06 dev/risk_scores_consortium_tests.Rmd
Work on risk scores

---
title: "Collaboration on risk scores"
output: html_notebook
---

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)

devtools::load_all()
library(tidyverse)
library(patchwork)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
```

Get marker counts

```{r}
relevant_markers <- c("lactate_dehydrogenase", "d_dimer", "platelet_count", "neutrophile_count", "lymphocyte_count", "AST")

subset_data <- data$marker_data %>% 
  filter(marker %in% relevant_markers & !is.na(value)) %>%
  group_by(marker, patient_id) %>% summarise(present = TRUE, .groups = "drop") %>%
  pivot_wider(names_from = "marker", values_from = "present") %>%
  mutate(across(all_of(relevant_markers), ~ if_else(is.na(.x), 0L, 1L)))

UpSetR::upset(subset_data %>% as.data.frame(), sets = relevant_markers)
```

No GFR (no Wang lab)