### 2021-06-03 other_outputs/fpv_metaanalysis.Rmd
Computing data for FPV metaanalysis

---
title: "Favipiravir - data for metaanalysis"
output: html_notebook
---

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8)

devtools::load_all()
library(tidyverse)
library(patchwork)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- data$marker_data_wide
```

```{r}
fpv_hospital_id <- wide %>% filter(took_favipiravir) %>% pull(hospital_id) %>% unique()

non_censored_patients <- data$patient_data %>% filter(outcome %in% c("Death", "Discharged"))
data_fpv <- filter_complete_data(data, 
                                 hospital_id == fpv_hospital_id & 
                                 patient_id %in% non_censored_patients$patient_id)
wide_fpv <- data_fpv$marker_data_wide
```

```{r}
data_fpv$marker_data %>% group_by(marker, patient_id) %>%
  summarise(n_days = n(), units = unique(unit)) %>%
  summarise(n_patients = n(), median_days = median(n_days), units = unique(units)) %>%
  arrange(desc(n_patients))
```

```{r}
markers_for_adverse_effects <- tribble(
  ~marker,            ~low_bound,      ~high_bound,
  "creatinin",        49,              90,
 # "CRP",
  "platelet_count",   150,             450,
  #"AST"
  "eosinophile_count",0,               0.5,
  "lymphocyte_count", 0.8,             4,
  "neutrophile_count",2,               7,
 # "albumin",          36,              45, # Removed, too few events/at risk
  #"pt_inr",                                # Removed - hard to put good bounds 
 # "procalcitonin",    0,               0.5 # Removed, too few events/at risk
)
```

```{r}
fpv_patients <- wide_fpv %>% 
  group_by(patient_id) %>%
  filter(any(took_favipiravir)) %>%
  summarise(min_day_fpv = min(day[took_favipiravir]), max_day_fpv = max(day[!is.na(favipiravir) & favipiravir > 0])) %>% 
  inner_join(data_fpv$patient_data, by = "patient_id") %>%
  select(patient_id, age, sex, min_day_fpv, max_day_fpv) %>%
  mutate(match_group = 1)

fpv_patients
```
```{r}
patients_matching <- data_fpv$patient_data %>%
  filter(!ever_favipiravir) %>%
  inner_join(data_fpv$marker_data, by = "patient_id") %>%
  group_by(patient_id, age, sex) %>%
  summarise(days_total.control = max(day), .groups = "drop") %>%
  mutate(match_group = 1)

potential_matches <- patients_matching %>%
  inner_join(fpv_patients, by = c("match_group"), suffix = c(".control", ".fpv")) %>%
  filter(days_total.control > min_day_fpv, abs(age.control - age.fpv) < 15)
```

```{r}
set.seed(48632485)
matches <- potential_matches %>%
  group_by(patient_id.control) %>%
#  filter(patient_id.fpv == sample(patient_id.fpv, size = 1, prob = exp(-abs(age.control - age.fpv))))
  filter(patient_id.fpv == sample(patient_id.fpv, size = 1, prob = 1/(1 + abs(age.control - age.fpv))))

matches_by_source <- matches %>% group_by(patient_id.fpv) %>% 
  summarise(count = n(), .groups = "drop")

if(nrow(matches_by_source) != nrow(fpv_patients)) {
  stop("Some FPV patients figure in no matching")
}
```

```{r}
groups_and_days_fpv <- 
  fpv_patients %>% select(patient_id, min_day_fpv, age, sex) %>%
  rename(exposure_day = min_day_fpv) %>%
  mutate(group = "FPV")

groups_and_days_control <- 
  matches %>% select(patient_id.control, min_day_fpv, age.control, sex.control) %>%
  rename(patient_id = patient_id.control, 
         age = age.control,
         sex = sex.control,
         exposure_day = min_day_fpv) %>%
  mutate(group = "Control")

groups_and_days <- rbind(groups_and_days_fpv, groups_and_days_control)

data_for_model <- data_fpv$marker_data %>%
  inner_join(groups_and_days, by = "patient_id") %>%
  inner_join(markers_for_adverse_effects, by = "marker") %>%
  ungroup() %>%
  mutate(outside_bounds = value < low_bound | value > high_bound) %>%
  group_by(patient_id, group, marker, age, sex) %>%
  summarise(at_risk = any(!outside_bounds[day <= exposure_day]) & any(!is.na(value[day > exposure_day & day < exposure_day + 10])),
            event = at_risk & any(outside_bounds[day >= exposure_day]), .groups = "drop")

data_for_model %>% 
  group_by(marker) %>%
  summarise(n_at_risk = sum(at_risk), n_events = sum(event), .groups = "drop")

data_for_model_all <- data_for_model %>% group_by(patient_id, age, sex, group) %>%
  summarise(at_risk = any(at_risk), event = any(event), .groups = "drop")

data_for_model_all %>% 
  summarise(n_at_risk = sum(at_risk), n_events = sum(event), .groups = "drop")

```

```{r}
fit_adverse_markers <- glm( event ~ age + sex + group, family = "binomial", data = data_for_model_all %>% filter(at_risk))

fpv_ci <-  confint(fit_adverse_markers)["groupFPV", ]
fpv_coeff <- summary(fit_adverse_markers)$coefficients["groupFPV", ]

other_adverse <- tibble::tribble(
  ~event, ~estimate, ~se, ~`lower95`, ~`upper95`,
  "Lab marker out of range", fpv_coeff["Estimate"], fpv_coeff["Std. Error"], fpv_ci[1], fpv_ci[2]
)

other_adverse
write_csv(other_adverse, here::here("local_temp_data", "fpv_meta", "lab_events.csv"))

```

```{r}
all_event_markers <- data_fpv$marker_data %>%
  inner_join(markers_for_adverse_effects, by = "marker") %>%
  mutate(outside_bounds = value < low_bound | value > high_bound) %>%
  group_by()

took_fpv <-  wide_fpv %>% select(patient_id, day,  took_favipiravir) %>%
  crossing(markers_for_adverse_effects)

all_event_markers_fpv <- took_fpv %>% 
  left_join(all_event_markers, by = c("patient_id", "day", "low_bound", "high_bound", "marker")) %>%
  mutate(group = if_else(took_favipiravir, "FPV", "Control"))

if(nrow(all_event_markers_fpv) != nrow(took_fpv)) {
  stop("bad join")
}
```

```{r}
all_res <- list()
for(exposure_day in 1:6) {

  all_events_risk <- all_event_markers_fpv %>%
    group_by(patient_id, marker) %>%
    summarise(at_risk = any(!outside_bounds[day <= exposure_day]) & any(!is.na(value[day > exposure_day & day < exposure_day + 10])),
              event = at_risk & any(outside_bounds[day >= exposure_day]), 
              group = group[day == exposure_day],
              .groups = "drop")

  by_marker <- all_events_risk %>%
    group_by(group, marker) %>%
    summarise(N = n(), at_risk = sum(at_risk, na.rm = TRUE), event = sum(event, na.rm = TRUE), .groups = "drop")

  any_marker <- all_events_risk %>%
    group_by(group) %>%
    summarise(N = length(unique(patient_id)), marker = "Any", at_risk = length(unique(patient_id[at_risk])), event = length(unique(patient_id[event])), .groups = "drop")

  all_res[[exposure_day]] <- rbind(any_marker, by_marker) %>% mutate(day = exposure_day)
}

do.call(rbind, all_res)

```

```{r}
all_event_markers_fpv %>% filter(is.na(group))

all_event_markers_fpv %>% ungroup() %>%
  group_by(day) %>%
  summarise(length(unique(patient_id)), sum(took_favipiravir), sum(is.na(took_favipiravir)), .groups = "drop")

dd <- all_event_markers_fpv %>%
  group_by(patient_id, marker) %>% summarise(at_risk = any(!outside_bounds[day <= exposure_day]) & any(!is.na(value[day > exposure_day & day < exposure_day + 10])), gr = if_else(took_favipiravir[day == exposure_day], "FPV", "Control"))

length(unique(dd$patient_id))
```

## Simple summaries

```{r}
table_fpv <- data_fpv$patient_data %>% 
  mutate(group = if_else(ever_favipiravir, "FPV", "Control")) %>% 
  group_by(group) %>% 
  summarise(
    N = n(),
    Male = sum(sex == "M"),
    `Age mean` = mean(age),
    `Age sd` = sd(age),
    `Age min` = min(age),
    `Age max` = max(age),
    `Admitted for Covid` = sum(admitted_for_covid),
    `Took hydroxychloroquine` = sum(ever_hcq),
    `Took azithromycin` = sum(ever_az),
    `Took dexamethasone` = sum(ever_dexamethasone),
    `Took remdesivir` = sum(ever_remdesivir),
    `Convalescent plasma` = sum(ever_convalescent_plasma),
    `Ischemic Heart Disease` = sum(ischemic_heart_disease),
    `Takes antihypertensives` = sum(has_hypertension_drugs),	
    `Heart Failure` = sum(heart_failure),
    COPD = sum(COPD),
    Asthma = sum(asthma),
    `Other lung disease` = sum(other_lung_disease),
    Diabetes = sum(diabetes),
    `Renal Disease` = sum(renal_disease),
    `Liver Disease` = sum(liver_disease),
    `Smoking` = sum(smoking),
    `BMI mean` = mean(BMI, na.rm = TRUE),
    `BMI sd` = sd(BMI, na.rm = TRUE),
    `Best supportive care` = sum(!is.na(best_supportive_care_from)),
    Deceased = sum(outcome == "Death"),
    `Required ventilation` = sum(worst_breathing_s >= "Ventilated"),
    `Required oxygen` = sum(worst_breathing_s >= "Oxygen"),
    Discharged = sum(outcome == "Discharged"),
    .groups = "drop"
  ) %>%
  as.data.frame()%>%
  column_to_rownames("group") %>%
  t() %>% as.data.frame() %>%
  rownames_to_column("characteristic")

table_fpv

write_csv(table_fpv, here::here("local_temp_data", "fpv_meta", "characteristics.csv"))

```